---
title: "Tutorial_1"
author: "Sven Panis"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Tutorial_1 file wrangles raw data from Experiment 1 of Panis and Schmidt (2016), and illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, when there is one independent variable.

# Load the libraries that we will be using. 

```{r load-pkg}
pkg <- c("tidyverse", "patchwork")

lapply(pkg, library, character.only = TRUE)
```

# Read in the raw data from Panis and Schmidt (2016).

```{r read-raw-data}
raw <- read_delim(file="Tutorial_1_descriptive_stats/data/DataExp1_6subjects.txt", delim="\t",col_names=T)
head(raw)
tail(raw)
glimpse(raw)
```

# Do some wrangling to get a simpler data set.

```{r wrangle-raw-data}
# remove practice trials
data <- raw %>% filter(practice==0) # 9240 trials: 6 participants x 70 blocks x 22 trials-per-block

# remove unnecessary columns
data <- data %>% select(-c(ss,practice,rd,form,task,primeduration,
                                             blankduration,maskduration,targetduration,
                                             void,target_dir,corresp,resperr))
# remove trials with no response info
sel = which(data$resp==0)
data <- data[-c(sel),] # 9189 rows left

# keep only the no-mask trials
data <- data %>% filter(mask_type == 0) %>%
                 select(-mask_type) # 2737 rows left, and 8 variables

summary(data)
# vp = Versuchsperson or subject
# bl = block number
# tr = trial number within block
# prime_type: no prime (0), consistent prime (1), inconsistent prime (2)
# resp = response left (1) or right (2)
# respac = accuracy
# rt = response time
# TrialNr = trial number within experiment
```

```{r save-data-file}
# save wrangled data file
write_csv(data, "Tutorial_1_descriptive_stats/data/DataExp1_6subjects_wrangled.csv")
```

# Read in simpler data file, change variable names, and create factors.

```{r read-wrangled-data}
data_wr <- read_csv("Tutorial_1_descriptive_stats/data/DataExp1_6subjects_wrangled.csv")
head(data_wr)

# change relevant variable names (to pid, trial, condition, rt, and acc)
colnames(data_wr) <- c("pid","bl","tr","condition","resp","acc","rt","trial") 
# pid to indicate column with unique person IDs
# trial must be a unique trial number per subject
# condition must be a factor indicating the levels (1,2,...) and labels of the single independent variable 
# rt to indicate column with response times (in ms)
# acc to indicate column with accuracies (1/0)

# create factor for condition (prime type)
data_wr <- data_wr %>% mutate(condition = condition + 1,
                              condition = factor(condition, levels = c(1,2,3), labels = c("blank","congruent","incongruent")))
summary(data_wr)
```

# Apply functional programming to set up life tables, and plot the discrete-time functions (h(t), S(t), and ca(t)).

Load the functions. Make sure you specify the same levels and labels for factor condition in the function plot_eha (line 205) as above (line 83).

```{r load-functions}
# function censor creates censored observations and discrete response times (drt), given a user-defined censoring time (timeout) and bin width in ms
censor <- function(df, timeout, bin_width){
  if(!(timeout %% bin_width == 0)){
    return("The censoring time must be a multiple of the bin width!")
  }
  if(timeout < 0 | bin_width < 0){
    return("Both timeout and bin_width must be larger than 0!")
  }
  df %>% mutate(right_censored = 0,
                rtc = ifelse(rt > timeout, timeout, rt) %>% round(digits=2), # censored response times
                right_censored = ifelse(rtc == timeout,1,right_censored), # 1 = right censored observation, 0 = observed rt
                drt = ceiling(rtc/bin_width), # drt = discrete response time
                cens_time = timeout, bin_width = bin_width) # save both user-defined parameters for plotting
}

# function ptb creates a person-trial-bin oriented data set
ptb <- function(df){
  df %>% uncount(weights = drt) %>% # create drt rows per trial
         group_by(trial) %>% 
         mutate(period = 1:n()) %>% # create time bin (or time period) ranks within each trial
         mutate(event = if_else(period == max(period) & right_censored == 0, 1, 0)) %>% # event = 1 indicates response occurrence
         ungroup()
}

# function setup_lt sets up a life table for each level of condition (1 independent variable)
setup_lt <- function(ptb){
  ptb %>% mutate(event = str_c("event", event)) %>%
          group_by(condition,period) %>% 
          count(event) %>% 
          ungroup() %>% 
          pivot_wider(names_from = event,
                      values_from = n) %>% 
          mutate(event0 = ifelse(is.na(event0),0,event0), # replace NA with 0
                 event1 = ifelse(is.na(event1),0,event1),
                 risk_set = event0 + event1) %>% # define the risk set
          mutate(hazard = (event1 / risk_set) %>% round(digits = 3)) %>% # calculate hazard estimate
          mutate(se_haz = sqrt((hazard * (1 - hazard)) / risk_set) %>% round(digits = 4)) %>% # standard error for hazard
          group_by(condition) %>%
          mutate(survival = (cumprod(1-hazard)) %>% round(digits = 4), # calculate survival estimate
                 term     = (cumsum(hazard / (risk_set * (1 - hazard)))) %>% round(digits = 7), # intermediate calculation
                 se_surv  = (survival * sqrt(term)) %>% round(digits = 5)  ) %>% # Greenwood's (1926) approximation
          ungroup() 
}

# function calc_ca calculates the conditional accuracies 
calc_ca <- function(df){
  df %>% filter(right_censored==0) %>%
         group_by(condition,drt,cens_time,bin_width) %>%
         summarize(ca = mean(acc) %>% round(digits = 2),
                   n = n(),
                   .groups = 'drop') %>%
         ungroup() %>%
         mutate(period = drt,
                se_ca = sqrt((ca * (1-ca)) / n) %>% round(digits = 3)) %>%
         select(-drt)
}

# function join_lt_ca joins the conditional accuracies to the life tables
join_lt_ca <- function(df1,df2){df1 %>% left_join(df2, join_by(condition,period))}

# function extract_median is used to extract the S(t).50 quantiles (i.e., the estimated median RTs)
extract_median <- function(df){
  above_pct50 <- df %>% 
      group_by(condition) %>%
      filter(survival > .5) %>% 
      slice(n()) # take last row
  below_pct50 <- df %>% 
      group_by(condition) %>%
      filter(survival < .5) %>% 
      slice(1) # take first row
  # pull period above
  period_above <- pull(above_pct50, period)
  # pull survivor function values
  survival_above <- pull(above_pct50, survival)
  survival_below <- pull(below_pct50, survival)
  # estimate median by interpolation
  median_period <- period_above+((survival_above-.5)/(survival_above-survival_below))*((period_above+1)-period_above)
}

# function plot_eha plots the hazard, survivor, and ca(t) functions, when there is one independent variable
plot_eha <- function(df,subj,haz_yaxis){ # set the upper limit of the y-axis for the hazard functions to haz_yaxis == 1 when plotting the data of individual subjects (see line 244)
  library(patchwork)
  cutoff <- df %>% pull(cens_time) %>% max(na.rm=T)
  binsize <- df %>% pull(bin_width) %>% max(na.rm=T)
  median_period <- extract_median(df)
  n_conditions <- nlevels(df$condition)
  data_median <- c()
  for(i in 1:n_conditions){
    data_median <- append(data_median, c(median_period[i], median_period[i]))
  }
  
  data_medians <- tibble(period= data_median,
                         survival = rep(c(.5, 0),n_conditions),
                         condition = rep(1:n_conditions, each=2))
# plot the hazard functions
p1 <- df %>% ggplot(aes(x=period, color=condition, group=condition)) +
  geom_line(aes(y=hazard)) +
  geom_point(aes(y=hazard), size=1) + labs(color="Condition") +
  geom_linerange(aes(ymin=hazard-se_haz, ymax=hazard+se_haz), show.legend = F) +
  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
                     limits = c(0,cutoff/binsize)) +
  scale_y_continuous(limits = c(0,haz_yaxis)) +
  labs(x="", y="h(t)", title = paste("Subject ", subj)) +
  theme(legend.background = element_rect(fill = "transparent"),
        panel.grid = element_blank(),
        legend.position = "top")
# plot the survivor functions
p2 <-df %>%
  ggplot(aes(x=period, color=condition, group=condition)) +
  geom_line(aes(y=survival), show.legend = F) +
  geom_point(aes(y=survival), size=1, show.legend = F) +
  geom_linerange(aes(ymin=survival-se_surv, ymax=survival+se_surv), show.legend = F) +
  # add vertical lines at the median RTs in the plot of the survivor functions using geom_path(). Make sure you apply the same levels and labels for the factor condition as above on line 83!
  geom_path(aes(x=period, y=survival, color=factor(condition, levels =c(1,2,3),labels=c("blank","congruent","incongruent"))),
            data = data_medians, 
            linetype = 3, show.legend = F) +
  
  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
                     limits=c(0,cutoff/binsize)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="", y="S(t)",
       colour="Condition") +
  theme(panel.grid = element_blank())
# plot the conditional accuracy functions
p3 <-df %>%
  ggplot(aes(x=period, color=condition, group=condition)) +
  geom_line(aes(y=ca), show.legend = F) +
  geom_point(aes(y=ca), size=1, show.legend = F) +
  geom_linerange(aes(ymin=ca-se_ca, ymax=ca+se_ca), show.legend = F) +
  scale_x_continuous(breaks = c(0,1:(cutoff/binsize)), labels=c(0,1:(cutoff/binsize)*binsize),
                     limits=c(0,cutoff/binsize)) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x="Time bin t's endpoint (ms)", y="ca(t)",
       colour="Condition") +
  theme(panel.grid = element_blank())

p1/p2/p3
}
```

Apply the functions

```{r apply-functions}
# create a dataframe with two columns (pid and data); the latter contains a list of nested data frames
data_nested <- data_wr %>% group_nest(pid)
# apply the functions using map() and map2()
data_final <- data_nested %>% 
                     mutate(censored  = map(data, censor, 600, 40)) %>%   # ! user input: censoring time, and bin width
                     mutate(ptb_data  = map(censored, ptb)) %>%           # create person-trial-bin dataset
                     mutate(lifetable = map(ptb_data, setup_lt)) %>%      # create life tables without ca
                     mutate(condacc   = map(censored, calc_ca)) %>%       # calculate ca
                     mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%    # create life tables with ca
                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha,1))  # create plots 
```

# Extract info, save plot for each subject, save a lifetable, create and save the person-trial-bin oriented data file for hazard modeling, and a data file for ca(t) modeling.

```{r extract-info}
# view original data (column 2) for subject 1:
pluck(data_final, 2, 1) # pluck() ignores the round(digits=2) function
print(pluck(data_final, 2, 1), n=20)

# view person-trial-bin data set (column 4) for subject 1:
pluck(data_final,4,1)
print(pluck(data_final,4,1),n=30)

# view lifetable (column 7) for each condition for subject 1
pluck(data_final, 7, 1) %>% filter(condition=="blank")
pluck(data_final, 7, 1) %>% filter(condition=="congruent")
pluck(data_final, 7, 1) %>% filter(condition=="incongruent")

# view plot (column 8) for subject 1. 
pluck(data_final, 8, 1)

# view the median RTs in a tibble (median_tb)
median_bin <- map(data_final$lifetable_ca, extract_median)
binsize    <- data_final$censored[[1]] %>% pull(bin_width) %>% max(na.rm=T)
median_rt  <- modify_depth(median_bin,1,~.x*binsize)
median_tb  <- map(median_rt, ~as_tibble(., rownames = "condition")) %>% list_rbind(names_to = "pid")
median_tb
```

```{r save-plots}
# save the plot for each subject
map2(paste0("Tutorial_1_descriptive_stats/figures/Plot_for_subject", data_final$pid, "_PanisSchmidt.png"), data_final$plot, ggsave, width = 6, height = 8, dpi = 600)
```

```{r create-single-lifetable}
# get the binsize in ms
binsize    <- data_final$censored[[1]] %>% pull(bin_width) %>% max(na.rm=T)
# extract the lifetable (column 7 in data_final) for condition 1 (neutral) for subject 6
lifetable_blank_s6 <- data_final[[7]][[6]] %>% 
                        filter(condition=="blank") %>% 
                        mutate(bin = period*binsize,
                               events = event1) %>%
                        select(bin,risk_set,events,hazard,se_haz,survival,se_surv,ca,se_ca) 
# get riskset for first bin
riskset <- lifetable_blank_s6 %>% select(risk_set) %>% slice(1) %>% pull()

# create a row for time point 0
row0 <- c(bin=0,risk_set=riskset,events=0,hazard=0,se_haz=0,survival=1,se_surv=0,ca=NA,se_ca=NA)
# add new row
lifetable_blank_s6_complete <- lifetable_blank_s6 %>%
                                 add_row(!!! row0) %>%
                                 arrange(bin)
lifetable_blank_s6_complete
```

```{r save-single-lifetable}
write_csv(lifetable_blank_s6_complete, file="Tutorial_1_descriptive_stats/tables/lifetable_neutral_s6.csv")
```

```{r create-file-hazard-modeling}
# select the person-trial-bin data set for each subject, unnest them, and select relevant variables.
input_modeling_hazard <- data_final %>% 
                         select(pid, ptb_data) %>%  unnest(ptb_data) %>%
                         select(pid,bl,tr,trial,condition,period,event)
```

```{r save-file-hazard-modeling}
write_csv(input_modeling_hazard, file="Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
```

```{r create-file-ca-modeling}
# select the censored data set for each subject, unnest them, discard trials that are censored (RT > 600 or no response detected before the deadline), and select relevant variables. 
input_modeling_ca <- data_final %>%
                     select(pid, censored) %>% unnest(censored) %>%
                     filter(right_censored == 0) %>% # only keep uncensored trials
                     select(pid,bl,tr,trial,condition,acc,drt)
```

```{r save-file-ca-modeling}
write_csv(input_modeling_ca, file="Tutorial_1_descriptive_stats/data/inputfile_ca_modeling.csv")
```

# Optional: Create single plot for data aggregated across subjects. This is only recommended when all subjects show the same qualitative patterns.

```{r aggregated-data-plot}
# create single subject number, and unique trial number
data_agg <- data_wr %>% mutate(oldtrial=trial,
                               oldpid=pid,
                               trial = 1:n(),
                               pid = 1)
# apply functions
data_agg_nested <- data_agg %>% group_nest(pid)
data_agg_final <- data_agg_nested %>% 
                     mutate(censored  = map(data, censor, 600, 40)) %>%   # ! user input: censoring time, and bin width
                     mutate(ptb_data  = map(censored, ptb)) %>%           # create person-trial-bin dataset
                     mutate(lifetable = map(ptb_data, setup_lt)) %>%      # create life tables without ca
                     mutate(condacc   = map(censored, calc_ca)) %>%       # calculate ca
                     mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%    # create life tables with ca
                     mutate(plot      = map2(.x = lifetable_ca, .y = pid, plot_eha,1))  # create plots 
# view plot
pluck(data_final, 8, 1)
```

