---
title: "packages"
author: "Sven Panis"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file installs packages.

There is no need to do this after the first installation, as it should all be 
controlled by renv(). 

This file is just for Sven's benefit, so that he can see how he installed them.

# install the libraries that we will be using #

## install ##

As I first created a public repo on Github (sven-panis/Tutorial_Event_History_Analysis) which I then cloned to a new R project (RStudo: File->New Project->Version Control->Git), I need to install and initialize renv myself.

```{r}
install.packages("renv")
renv::init()
```

I then closed the project, reopened it, and added the following two lines to .Rprofile (so that git will not ignore the renv.lock file, and renv::snapshot() will recored everything):
options(renv.settings.vcs.ignore.local = "FALSE") 
renv::settings$snapshot.type("all") 


Next, general packages for data wrangling and visualisation used in various scripts.

```{r}
install.packages(c("tidyverse", "RColorBrewer", "patchwork", "ggridges"))
```

Next, packages for simulation (Figures 1, 2, and 3)

```{r}
install.packages(c("faux","ggpmisc","truncnorm"))
```


Next, packages for model building in brms used mainly in models.Rmd and effects.Rmd
files.

```{r}
install.packages("remotes")
remotes::install_github("stan-dev/cmdstanr")
library(cmdstanr)
install_cmdstan(cores=11) 
check_cmdstan_toolchain() 
## in the above, set cores correctly for your machine and without a folder location
## specified it just uses the default folder location. The default is fine for most
## setups though. To set the folder, include: dir = "/path/.cmdstan"

## you may also need to configure the toolchain, if you use a mac like I do:
## https://github.com/stan-dev/rstan/wiki/Configuring-C---Toolchain-for-Mac
## https://mac.thecoatlessprofessor.com/macrtools/


install.packages("devtools")
devtools::install_github("jmgirard/standist")
# 
install.packages(c("brms", "tidybayes", "bayesplot", "future"))

install.packages("broom.mixed")
```

To build regression models in other packages such as lme4:

```{r}
install.packages("lme4", "nlme")
```

Next, packages for papaja and manuscript writing

```{r install-pkg}
# install.packages(c("devtools", "remotes"))
# 
devtools::install_github("crsh/papaja")
devtools::install_github("crsh/citr")
# 
install.packages("tinytex")
tinytex::install_tinytex()
# 
# # and now some useful addins for writing .Rmd from R Studio word count
devtools::install_github("benmarwick/wordcountaddin", type = "source", dependencies = TRUE)
# 
# # remedy for shortcuts in markdown writing
remotes::install_github("ThinkR-open/remedy")
# 
# # grammar checking - cool...
devtools::install_github("ropenscilabs/gramr")
```

Finally, use renv() to take a snapshot of installed packages and update the lock.file 

```{r snapshot-renv}
# take a snapshot and update the lock.file
renv::snapshot() # this is only necessary when new packages or installed or packages are updated.
```