---
title: "Tutorial_4"
author: "Rich"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=T)
```

Tutorial 4 shows one way to plan the design of a future experiment using EHA.

# 1. Load the libraries that we will be using #

```{r load-pkg, results='hide'}
pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
         "brms", "tidybayes", "bayesplot", "future", "parallel", "VGAM", "faux",
         "rstan", "osfr", "here")

lapply(pkg, library, character.only = TRUE)
```

Set options. 

```{r set-options, results='hide'}
options(brms.backend = "cmdstanr",
        mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

rstan_options(auto_write = TRUE)

supportsMulticore()
detectCores()
```

```{r check-info, results='hide'}
# packageVersion("cmdstanr")
# devtools::session_info("rstan")
```

theme settings for ggplot

```{r plot-settings, results='hide'}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 22, face = "bold"), 
          title = element_text(size = 22, face = "bold"),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# 2. A brief outline of the workflow #

## 2.1 - Main background sources ##

The basic approach is based upon two main sources:

The basic structure and code follows the examples outlined by Solomon Kurz in his
'power' blog posts and Lisa Debruine's faux{} package. 

For Solomon's blog posts, see here: [https://solomonkurz.netlify.app/tags/power/](https://solomonkurz.netlify.app/tags/power/)

For Lisa's faux package, see here: [https://debruine.github.io/faux/](https://debruine.github.io/faux/).
The vignettes provided by the faux package are particularly helpful and worth following before you take a deep dive into the below code. 
There is also an associated [journal article](https://doi.org/10.1177/2515245920965119), which is worth reading. 

And for some simpler worked examples of the general workflow (i.e., simpler than using EHA), then take a look at these examples by Rich, which show how you might simulate multi-level data for a range of factorial designs: [https://github.com/rich-ramsey/sim_demo](https://github.com/rich-ramsey/sim_demo)

## 2.2 - Basic workflow ##

The basic workflow is as follows:

1. Fit a regression model to existing data.
2. Use the regression model parameters to simulate new data.
3. Write a function to create 1000s of datasets and vary parameters of interest (e.g., sample size, trial count, effect size).
4. Summarise the simulated data to estimate likely power or precision of the various research designs.

## 2.3 - Exceptions to the normal workflow ##

The basic workflow above is a modified version of that recommended by Solomon and Lisa.
In a typical workflow, once the datasets are simulated, a regression model would be fit to each dataset.
Then, the parameters from each model would be summarised to make a judgment about the likely precision or power of the design.
However, in the case of some of the models in this tutorial, which can take several hours to build, we are not able to easily build 1000 models per variation in parameter setting, such as sample size and/or trial count, as that would take an unreasonably long time on a desktop machine.
Instead, we choose to take a simpler approach and summarise the data per bin and condition, rather than the parameter estimates from the model.
Although this is not as satisfying, we feel it can still be a useful guide to set expectations. 
To further save time and computational resources, for the purposes of this planning tutorial, we also choose to use a smaller dataset than was used in the earlier tutorials. 
The details on this simpler dataset and modelling procedure are provided below.

# 3. Read in the data #

Here, we again use data form Panis and Schmidt (2016), but to make things simpler, we focus on only 6 time bins (time bins 4-9) and two prime conditions (congruent and incongruent).

## read in and wrangle the raw data ##

```{r load-data}
data <- read_csv("Tutorial_4_planning/data/tidyppp_Exp1_n6.csv") %>%
  ## no mask condition only, neutral prime removed, time bins 4-9 only
  filter(no_ma == 1, no_pr < 1, timebin %in% c(4:9)) %>% 
  ## remove unnecessary columns
  select(-c(bl, target_dir, corresp, resp, respac, rel_ma, irr_ma, ran_ma, no_ma)) %>%
  ## rename the error column
  rename(error = resperr) %>% 
  ## create a new variable and some factors
  mutate(prime = if_else(con_pr == 1, "con", "inc"),
         prime = factor(prime,
                        levels = c("con", "inc")),
         timebin = factor(timebin)) %>% 
  ## simplify the data by selecting a few variables only
  select(pid, timebin, prime, outcome)
head(data)
str(data)
```

data check

```{r check-conds}
data %>% 
  distinct(timebin, prime)
```

## create summary data ##

summary data

```{r calculate-summary-data}
data_pid <- data %>% 
  group_by(pid, timebin, prime) %>% 
  summarise(outcome = mean(outcome, na.rm = TRUE))
head(data_pid)

data_group <- data %>% 
  group_by(timebin, prime) %>% 
  summarise(outcome = mean(outcome, na.rm = TRUE))
head(data_group)
```

quick plot

```{r quick-summary-plot}
p3.1 <- ggplot(data_group, aes(x=timebin, y = outcome, colour = prime)) +
   geom_line(aes(group = prime, linetype = prime)) + geom_point() +
   geom_jitter(data=data_pid, alpha = 0.5, width = 0.1, height = 0) +
   scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
   scale_x_discrete(labels = c(1:6)) +
   scale_fill_brewer(palette = "Dark2") +
   scale_colour_brewer(palette = "Dark2") +
   labs(title = "Panis & Schmidt (2016) (N=6)",
        x = "time bins")
p3.1
```

check the number of trials

```{r check-trials}
tally <- data %>% 
  group_by(pid, prime, timebin) %>%
  summarise(n=n(),
            sum=sum(outcome),
            .groups = "drop")
tally
```

# 4. Build an initial model #

In general, it can sometimes be a lot simpler to use an index coding approach to build regression models. 

see this link for some background on an index coding approach:
https://bookdown.org/content/4857/the-many-variables-the-spurious-waffles.html#many-categories.

and then maybe specify the interaction term like this...
https://discourse.mc-stan.org/t/specifying-formulates-for-interaction-between-categorical-variables-with-the-index-coding-approach-in-brms/29449/3

And in the context of data simulation, index coding can be particularly useful, because it may be often easier and more intuitive to set condition means, rather than more complicated regression parameters, such as interaction terms.

The approach here is to build an initial model using an index coding approach, but with a subset of the data that was used in prior tutorials.

Because we are building Bayesian regression models, which generate a posterior distribution per condition of interest in our index coding approach, we can simply calculate contrasts of interests and associated precision or interval estimates (95% quantile intervals, for example) to address any questions we may have about comparisons between conditions, such as incongruent vs congruent in a particular time bin. 

## formula ##

with varying intercept and slope for prime by pid

```{r model-formula}
formula = bf(outcome ~ 0 + timebin:prime +
               (0 + timebin:prime | pid))
```

## check the priors available ##

```{r priors}
get_prior(formula,
          data = data, family = bernoulli(link = "cloglog"))
```

## visualise priors ##

Using an index coding approach, each condition has a prior for the mean of that condition.
And as we know from plot 3.1 above, the condition averages across the six time bins selected range from near zero on the hazard probability scale to 0.50. 
And in principle, the more general case could involve any hazard value from zero to 1.
Therefore, if we try to stick to a weakly informative approach to priors [see here for more details](https://github.com/stan-dev/stan/wiki/prior-choice-recommendations), then priors for the cloglog values should cover a fairly broad range.

For a look at values across the cloglog scale and how it compares to the logit scale, see [here](https://towardsdatascience.com/a-gentle-introduction-to-complementary-log-log-regression-8ac3c5c1cd83).

Ok, so let's take a look at some cloglog values at the lower and upper end of the probability scale.

```{r cloglog-vals}
cloglog_vals = clogloglink(c(0.01,0.99))
cloglog_vals

# -4.600149  1.527180
```

and now take a look at some distributions

```{r visualise-priors}
visualize("normal(0, 2)", "normal(0, 1)", "normal(0, 0.5)",
          xlim = c(-4, 4))
```

(0,2) looks like it covers all of our likely values, even if it is too broad.

For simplicity, we use these normal(0,2) priors for the mean per condition.
However, if we were doing this for real, we would probably choose different priors that are similar to ones that were discussed in Tutorial_2a, as well as in Supplementary Materials.

## set priors ##

We set the prior for the mean as (normal(0,2)) and then we use some default recommendations from McElreath 2020 for the rest.

```{r set-priors}
priors <- c(
  set_prior("normal(0, 2)", class = "b"),
  set_prior("normal(0, 1)", class = "sd"),
  set_prior("lkj(2)", class = "cor")
)
```

## run the model ##

This model takes ~17 minutes to build on a 2020 MacBook Pro (2 GHz Quad-Core Intel Core i5).
For this tutorial, this chunk is skipped to save time. 
Change eval=TRUE in the code chunk to run this model.

```{r build-model, eval=FALSE}
plan(multicore)
bi <- brm(formula = formula,
        data = data, family = bernoulli(link = "cloglog"),
        prior = priors,
        iter = 2000, warmup = 1000, cores = 8, chains = 4,
        control = list(adapt_delta = 0.95),
        save_pars = save_pars(all=TRUE),
        seed = 123,
        init = 0.01,
        file = "Tutorial_4_planning/models/bi")
summary(bi)
```

The model built without any concerns, errors or warnings.

At this point, you would normally do the following:

- check the model by looking at the convergence of chains and other model diagnostics. 
We'll skip this step for now, just to move on to the main focus of planning and data simulation. 

- Summarise and visualise the parameter estimates and/or posterior predictions.

Below, we quickly summarise fixed effects from the model in cloglog values, and we also convert them to hazard values, which tend to be far easier to interpret.

## visualise fixed effect parameters from model bi ##

read in the existing model object (if you did not build the model yourself)

```{r load-model-object}
bi <- readRDS("Tutorial_4_planning/models/bi.rds")
summary(bi)
```

wrangle the posterior distribution

```{r wrangle-posterior}
## take the posterior draws
post <- as_draws_df(bi) %>% 
  select(-lp__) %>% 
  as_tibble()

## create a summary, and use medians with robust = TRUE
post_summary <- posterior_summary(bi, robust = TRUE)

## just look at the fixed effects
post_qi_b <- post %>%
  select(starts_with("b_")) %>%
  pivot_longer(everything()) %>%
  group_by(name) %>% 
  median_qi(value)
head(post_qi_b)
```

a quick plot using posterior samples and tidybayes

```{r plot-fixed}
tidy_fixed <- post %>% 
  select(starts_with("b_"), .chain, .iteration, .draw) %>% # select and rename in simpler labels.
  rename(chain=.chain, iter=.iteration, draw=.draw) %>% 
  pivot_longer(-c(chain, draw, iter)) %>% # move from wide to long
  mutate(key = factor(name, levels=unique(name)),
         bin = rep(1:6, length.out = 48000),
         prime = rep(c("con", "inc"), each = 6, length.out = 48000),
         bin=factor(bin),
         prime=factor(prime))
head(tidy_fixed)
tail(tidy_fixed)

check.labels <- tidy_fixed %>% 
  distinct(key, bin, prime) 
check.labels

# plot
p4.1 <- ggplot(tidy_fixed, aes(x = bin, y = value, fill=prime)) +  
  stat_halfeye(alpha=0.7) +
  labs(title = "Model bi cloglog",
       x = "time bin", y = "cloglog") +
  scale_fill_brewer(palette = "Dark2") 
p4.1
```

```{r eval=F}
ggsave ("Tutorial_4_planning/figures/index_cloglog.jpeg",
        width = 6, height = 4, dpi = 800)
```

calculate hazards per condition per bin in the posterior dist

```{r calculate-bi-hazard}
tidy_haz <- tidy_fixed %>%
  select(chain, iter, draw, bin, prime, value) %>% 
  mutate(hazard = exp(value))
head(tidy_haz)
```

plot hazards

```{r plot-bi-hazards}
p4.2 <- ggplot(tidy_haz, aes(x = bin, y = hazard,
                                    fill=prime)) +  
  stat_halfeye(alpha=0.7) +
  labs(title = "hazard estimates",
       x = "time bin", y = "hazard") +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) 
p4.2
```

```{r eval=F}
ggsave ("Tutorial_4_planning/figures/index_hazard.jpeg",
        width = 6, height = 4, dpi = 800)
```

warning about removing rows. 

# 5. Simulate a single dataset #

Note: For multi-level data, data simulation can be quite a convoluted process, and we therefore strongly recommend reading the power blogs by Solomon Kurz and working through the faux package [vignettes](https://debruine.github.io/faux/index.html), before getting stuck into the below code.

read in a prior model, if not already loaded.

```{r load-model-object-sim, eval=FALSE}
bi <- readRDS("Tutorial_4_planning/models/bi.rds")
summary(bi)
```

Define some parameters

```{r define-sim-parameters}
# define parameters
# specify some features of the design
subj_n = 10  # number of subjects
rep_n = 200 # number of trial repeats 
cond_n = 2 # number of conditions
bin_n = 6 # number of time bins

## set fixed effects
## as a reminder
fixef <- as_tibble(fixef(bi), rownames = "term") %>%
  mutate(across(where(is.double), \(x) round(x, 2)))
fixef

## b1 to b6 = time bin 1 to 6.
## c = con
b1c = -4.66
b2c = -3.24
b3c = -2.32
b4c = -1.49
b5c = -1.21
b6c = -0.60

## i = inc
b1i = -4.37
b2i = -3.21
b3i = -2.37
b4i = -2.27
b5i = -3.04
b6i = -2.45

## set varying effects by pid
## as a reminder
varcor <- VarCorr(bi)
glimpse(varcor)
str(varcor)

## extract sd
sd <- as_tibble(varcor$pid$sd, rownames = "term") %>%
  mutate(across(where(is.double), \(x) round(x, 2)))
sd

u1c_sd = 1.06   # 
u2c_sd = 1.34   # 
u3c_sd = 0.60   # 
u4c_sd = 0.47   # 
u5c_sd = 0.58   #
u6c_sd = 0.72   # 

u1i_sd = 1.17   # 
u2i_sd = 1.05   # 
u3i_sd = 0.91   # 
u4i_sd = 1.47   # 
u5i_sd = 1.34   #
u6i_sd = 0.62   # 
```

make a correlation matrix 

```{r define-cor-mat}
# varcor <- VarCorr(b3c_out)
# varcor
# str(varcor)

## take each one
sd <- varcor$pid$sd
cor <- varcor$pid$cor
cov <- varcor$pid$cov

## make it tidy
cor_mat_tidy <- as_tibble(cor, rownames = "term") %>% ## 
  rename_with(~str_replace_all(.x, '\\.', '_')) %>% 
  pivot_longer(-term,
               values_to = "value",
               names_to = c("parameter", "term2"),
               names_pattern = "([a-zA-Z_\\d?]+)_([a-zA-Z:\\d?]+)") %>% ## pivot longer and use names_pattern to select the parts of the column names to separate on. the above is regex for each bit between the separator ("-"). e.g., ([a-zA-Z]+\\d+). this looks for letters and digits before the first "_". This (\\d?) looks for the possibility of digits.
  ## the rest of this code just turns the variables into what we want
  filter(parameter == "Estimate") %>%
  select(-parameter) %>% 
  pivot_wider(names_from = "term2",
              values_from = "value") %>% 
  select(-term)
head(cor_mat_tidy)
str(cor_mat_tidy)

## check with the model summary
# summary(bi)

## remove names and make a matrix as faux{} likes it like that
## full cor mat
cor_mat <- unname(as.matrix(cor_mat_tidy))
cor_mat
```

setup the data structure

```{r setup-sim-d1}
# make it reproducible
set.seed(1)

d1 <- add_random(subj = subj_n, rep = rep_n) %>%
  add_within("subj", condition = c("cond1", "cond2")) %>%
  add_contrast("condition", "treatment", add_cols = TRUE, 
               colnames = c("cond")) %>%
  add_within("rep", bin = 1:bin_n) %>%
  ### create new conditions here that respect the twelve levels b1c, b1i etc.
  mutate(con = if_else(condition == "cond1", 1, 0),
         inc = if_else(condition == "cond2", 1, 0)) %>% 
  mutate(bin1 = if_else(bin == 1, 1, 0),
         bin2 = if_else(bin == 2, 1, 0),
         bin3 = if_else(bin == 3, 1, 0),
         bin4 = if_else(bin == 4, 1, 0),
         bin5 = if_else(bin == 5, 1, 0),
         bin6 = if_else(bin == 6, 1, 0)) %>% 
  # add random effects 
  add_ranef("subj", u1c = u1c_sd, u2c = u2c_sd, u3c = u3c_sd, u4c = u4c_sd,
            u5c = u5c_sd, u6c = u6c_sd,
            u1i = u1i_sd, u2i = u2i_sd, u3i = u3i_sd, u4i = u4i_sd,
            u5i = u5i_sd, u6i = u6i_sd, 
            .cors = cor_mat) %>% 
  # calculate logit
  mutate(cloglog = (b1c + u1c) * con * bin1 + (b2c + u2c) * con * bin2 + 
           (b3c + u3c) * con * bin3 + (b4c + u4c) * con * bin4  + 
           (b5c + u5c) * con * bin5 + (b6c + u6c) * con * bin6 +
           (b1i + u1i) * inc * bin1 + (b2i + u2i) * inc * bin2 + 
           (b3i + u3i) * inc * bin3 + (b4i + u4i) * inc * bin4  + 
           (b5i + u5i) * inc * bin5 + (b6i + u6i) * inc * bin6) %>% 
  # calculate inverse cloglog
  mutate(prob = clogloglink(cloglog, inverse = TRUE)) %>%  
  # calculate event
  mutate(event = rbinom(n(), 1, prob)) 
head(d1)
str(d1)
summary(d1)
glimpse(d1)

## save initial data
# write_csv(d1, "Tutorial_4_planning/data/d1.csv") #
```

summarise the data produced so far

```{r summarise-sim-d1}
d1_pid <- d1 %>%
  group_by(subj, condition, bin) %>%
  summarise(
    across(cloglog:event, \(x) mean(x, na.rm = TRUE)),
    .groups = "drop"
    )
d1_pid  

d1_group <- d1 %>%
  group_by(condition, bin) %>%
  summarise(
    across(cloglog:event, \(x) mean(x, na.rm = TRUE)),
    .groups = "drop"
    )
d1_group
```

a quick plot

cloglog

```{r plot-d1-clog}
p5.1 <- ggplot(d1_group, aes(x=bin, y = cloglog, colour = condition)) +
   geom_line(aes(group = condition)) + 
   geom_point() +
   geom_jitter(data=d1_pid, alpha = 0.5, width = 0.1, height = 0) +
   # scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
   scale_x_continuous(breaks = 1:bin_n, labels = 1:bin_n) +
   scale_fill_brewer(palette = "Dark2") +
   scale_colour_brewer(palette = "Dark2") +
   labs(title = "cloglog - simulation d1 (N=10)",
        x = "time bins")
p5.1
```

event prob

```{r plot-d1-prob}
p5.2 <- ggplot(d1_group, aes(x=bin, y = event, colour = condition)) +
   geom_line(aes(group = condition, linetype=condition)) + 
   geom_point() +
   geom_jitter(data=d1_pid, alpha = 0.5, width = 0.1, height = 0) +
   scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
   scale_x_continuous(breaks = 1:bin_n, labels = 1:bin_n) +
   scale_fill_brewer(palette = "Dark2") +
   scale_colour_brewer(palette = "Dark2") +
   labs(title = "event prob - simulation d1 (N=10)",
        x = "time bins")
p5.2
```

convert the data into outcome data (this next chunk can be incorporated into the data structure chunk above, I'm just doing separately here, so that you can see what is initially generated)

```{r calculate-outcome-d1}
d1_out <- d1 %>% 
  group_by(subj, rep, condition) %>%
  mutate(cumsum = cumsum(event),
         outcomel = event==1 & cumsum(event) < 2,
         outcomen = if_else(event==1 & cumsum(event) == 1, 1,
                     if_else(cumsum(event) >= 1, NA, 0))) %>% 
  drop_na(outcomen) %>% 
  ungroup()
head(d1_out)

## save outcome data
# write_csv(d1_out, "Tutorial_4_planning/data/d1_out.csv")
```

summarise the outcome data, which is produced based on the prob value, right?

```{r summary-d1-out}
d1_out_pid <- d1_out %>% 
  group_by(subj, condition, bin) %>% 
  summarise(outcome = mean(outcomen, na.rm = TRUE)) %>% 
  mutate(prime = if_else(condition == "cond1", "con", "inc"),
         prime = factor(prime))
d1_out_pid

d1_out_group <- d1_out %>% 
  group_by(condition, bin) %>% 
  summarise(outcome = mean(outcomen, na.rm = TRUE)) %>% 
  mutate(prime = if_else(condition == "cond1", "con", "inc"),
         prime = factor(prime))
d1_out_group
```

plot outcome using the prime variable to make it comparable to the raw data

```{r plot-d1-outcome}
p5.3 <- ggplot(d1_out_group, aes(x = bin, y = outcome, colour = prime)) +
  geom_line(aes(group = prime, linetype = prime)) + 
  geom_point() +
  geom_jitter(data=d1_out_pid, alpha=0.5, width = 0.1, height = 0) +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
  scale_x_continuous(breaks = 1:bin_n, labels = 1:bin_n) +
  scale_colour_brewer(palette = "Dark2") +
  labs(title = "simulation d1 (N=10)",
       x = "time bins",
       y = "outcome")
p5.3
```

compare the raw data to the simulated data for a single simulated dataset

```{r plot-raw-plus-d1}
p5.4 <- (p3.1 / p5.3) +
   plot_layout(guides = 'collect',
               axes = 'collect')
p5.4
```

```{r eval=F}
ggsave ("Tutorial_4_planning/figures/raw_vs_d1.jpeg",
        width = 8, height = 8, dpi=800)
```

ok, this looks reasonably sensible.

# 6. Think about effect sizes

take a look at avg hazards and hazard ratios at the group level in the original data

```{r plot-raw}
p3.1
```

this is the descriptive data per bin and condition for the hazard - probability of a response/event, given that it has not happened yet

this is the group level summary data plotted in dat_plot

```{r plot-raw-summary}
data_group
```

calculate the hazard ratio

```{r calculate-hazard}
hazard <- data_group %>% 
  pivot_wider(names_from = "prime",
              values_from = "outcome") %>% 
  mutate(diff = con-inc,
         ratio = inc/con)
hazard

# timebin     con    inc     diff ratio
#   <fct>     <dbl>  <dbl>    <dbl> <dbl>
# 1 4       0.00975 0.0139 -0.00416 1.43 
# 2 5       0.0478  0.0423  0.00551 0.885
# 3 6       0.0901  0.0913 -0.00121 1.01 
# 4 7       0.188   0.115   0.0732  0.611
# 5 8       0.252   0.0495  0.203   0.196
# 6 9       0.401   0.0848  0.316   0.211
```

key hazard and hazard ratio values of interest are as follows....

```{r key-hazards, eval=FALSE}
## bin4
## 0.61 - 39% reduction in hazard from con to inc
## 0.115/0.188

## bin5
## 0.20 - 80% reduction in hazard from con to inc
## 0.0495/0.252

## bin6
## 0.21 - 79% reduction in hazard from con to inc
## 0.401/0.0848
```

Ok, if we wanted to simulate a bunch of effect sizes, we could think about the following.

Let's just focus on bin 6 to keep things simple.
And keep in mind that we do need to think about the absolute hazard values, as well as the ratio.

So con is close to 0.1, so let's fix it there as a round number and calculate some deviations from it.

Given that the effect of interest was ~80% reduction in Sven's 2016 paper (Panis & Schmidt, 2016), let's play around with a range of effect sizes from e.g., 0.25, 0.5, 0.75.
These would translate to a 75% reduction, 50% reduction and 25% reduction from con to inc.

```{r sim1-bin6-effect-sizes}
bin6 <- tibble(
  effect_id = 1:3,
  inc = rep(0.1, times=3),
  ratio = c(0.25, 0.5, 0.75),
  con = inc/ratio,
  inc_clog = clogloglink(inc),
  con_clog = clogloglink(con)
)
bin6

#  effect_id   inc ratio   con inc_clog con_clog
#       <int> <dbl> <dbl> <dbl>    <dbl>    <dbl>
# 1         1   0.1  0.25 0.4      -2.25   -0.672
# 2         2   0.1  0.5  0.2      -2.25   -1.50 
# 3         3   0.1  0.75 0.133    -2.25   -1.94 
```

# 7. Create a function to iterate through and simulate many datatsets

create a function to simulate data
These values are from the initial model and data from Sven's paper, which we created above in section 4, but they can be modified when running the function, which we do below.

```{r index-sim-function}
index_sim <- function(subj_n = 10, rep_n = 200, bin_n = 6,  # these can be changed when calling the function
                b1c = -4.66, b2c = -3.24, b3c = -2.32, b4c = -1.49, b5c = -1.21, b6c = -0.6, 
                b1i = -4.37, b2i = -3.21, b3i = -2.37, b4i = -2.27, b5i = -3.04, b6i = -2.45, # fixed effects
                u1c_sd = 1.06, u2c_sd = 1.34, u3c_sd = 0.60, u4c_sd = 0.47, u5c_sd = 0.58, u6c_sd = 0.72, 
                u1i_sd = 1.17, u2i_sd = 1.05, u3i_sd = 0.91, u4i_sd = 1.47, u5i_sd = 1.34, u6i_sd = 0.62, # varying effects
                cors = cor_mat,   # correlations between effects
                ... # helps the function work with pmap() below
                ) {
  # set up data structure
  data <- add_random(subj = subj_n, rep = rep_n) %>%
  add_within("subj", condition = c("cond1", "cond2")) %>%
  add_contrast("condition", "treatment", add_cols = TRUE, 
               colnames = c("cond")) %>%
  add_within("rep", bin = 1:bin_n) %>%
  ### create new conditions here that respect the twelve levels b1c, b1i etc.
  mutate(con = if_else(condition == "cond1", 1, 0),
         inc = if_else(condition == "cond2", 1, 0)) %>% 
  mutate(bin1 = if_else(bin == 1, 1, 0),
         bin2 = if_else(bin == 2, 1, 0),
         bin3 = if_else(bin == 3, 1, 0),
         bin4 = if_else(bin == 4, 1, 0),
         bin5 = if_else(bin == 5, 1, 0),
         bin6 = if_else(bin == 6, 1, 0)) %>% 
  # add random effects 
  add_ranef("subj", u1c = u1c_sd, u2c = u2c_sd, u3c = u3c_sd, u4c = u4c_sd,
            u5c = u5c_sd, u6c = u6c_sd,
            u1i = u1i_sd, u2i = u2i_sd, u3i = u3i_sd, u4i = u4i_sd,
            u5i = u5i_sd, u6i = u6i_sd, 
            .cors = cor_mat) %>% 
  # calculate logit
  mutate(cloglog = (b1c + u1c) * con * bin1 + (b2c + u2c) * con * bin2 + 
           (b3c + u3c) * con * bin3 + (b4c + u4c) * con * bin4  + 
           (b5c + u5c) * con * bin5 + (b6c + u6c) * con * bin6 +
           (b1i + u1i) * inc * bin1 + (b2i + u2i) * inc * bin2 + 
           (b3i + u3i) * inc * bin3 + (b4i + u4i) * inc * bin4  + 
           (b5i + u5i) * inc * bin5 + (b6i + u6i) * inc * bin6) %>% 
  # calculate inverse cloglog
  mutate(prob = clogloglink(cloglog, inverse = TRUE)) %>%  
  # calculate event
  mutate(event = rbinom(n(), 1, prob)) %>% 
  group_by(subj, rep, condition) %>%
  mutate(cumsum = cumsum(event),
         outcomel = event==1 & cumsum(event) < 2,
         outcomen = if_else(event==1 & cumsum(event) == 1, 1,
                     if_else(cumsum(event) >= 1, NA, 0))) %>% 
  drop_na(outcomen) %>%
  ungroup()

  # glimpse(data) # only use this when testing the code
}
```

Here’s a quick example of how our function works. You can change these parameters
and create some different data.

```{r test-index-sim}
test_sim <- index_sim(subj_n = 10, rep_n = 200) # if you uncomment glimpse above,
# it will let you glimpse the data that's generated. this is useful for checking / testing code purposes.
```

create a function to summarise the simulated data

```{r index-summary-function}
index_summary <- function(df) {
  df %>% 
  group_by(subj, condition, bin) %>% 
  summarise(n = n(), 
            sum = sum(outcomen),
            mean_outcome = mean(outcomen, na.rm = TRUE),
            sd_outcome = sd(outcomen, na.rm = TRUE),
            sem = (sd_outcome/sqrt(length(unique(rep)))),
            .groups="drop") 
}
```

test it

```{r test-index-summary}
test_summary <- index_summary(test_sim)
head(test_summary)
```

# 8. Simulate data for a range of parameters

vary trial count per condition and effect size, keep pid fixed at N=10.
This will take some time, probably hours...go get a beer/coffee.
For the purposes of this tutorial, eval is set to false in the below chunks, to save time.

```{r run-sim1, eval=FALSE}
plan(multicore)
x1 <- crossing(
  exp = 1:1000, # number of experiment replicates
  subj_n = 10, # range of subject N
  rep_n = c(100, 200, 400),
  b6i = -2.25,
  b6c = c(-1.94, -1.50, -0.67)
) %>%
  mutate(d = pmap(., index_sim)) %>% 
  mutate(s = map(d, index_summary)) %>% 
  select(-d)
```

# 9. Summarise the simulated data

```{r summarise-sim1, eval=FALSE}
sx1 <- x1 %>%
  unnest(s) %>%
  mutate(exp=factor(exp),
         subj_n=factor(subj_n),
         rep_n=factor(rep_n),
         b6c = factor(b6c,
                      levels = c("-1.94", "-1.5", "-0.67"),
                      labels = c("25%", "50%", "75%")),
         subj=factor(subj),
         bin=factor(bin))
sx1
```

take a look

```{r sim1-quick-look, eval=FALSE}
head(sx1)
tail(sx1)

sx1 %>% 
  distinct(b6c)
```

save out a file 

```{r save-sim1, eval=FALSE}
## save the first simulation
write_csv(sx1, "Tutorial_4_planning/data/sim1/sim1_data.csv")
```

## calculate summary data ##

read in the file, if it is already computed.

If necessary, read it from the OSF to start with (because it is a large file). 
This step is only required to be done once.
After that, the file will be in your local directory and you can skip this chunk in the future.

```{r load-osf-sim1, eval=T}
## create a relevant folder, if it does not already exist
sim1_output_dir <- here("Tutorial_4_planning/data/sim1/")

if (!dir.exists(here(sim1_output_dir))) {
  dir.create(here(sim1_output_dir))
}

## read in the data from the OSF and store it in a relevant folder, if it does not already exist
sim1_file_name <- here("Tutorial_4_planning/data/sim1/sim1_data.csv")

if (!file.exists(here(sim1_file_name))) {
  osf_retrieve_node("3dbcs") %>% 
  osf_ls_files(pattern = "Tutorial_4_planning") %>%
  osf_ls_files(pattern = "data") %>%
  osf_ls_files(pattern = "sim1") %>%
  osf_ls_files(pattern = "sim1_data") %>%
  osf_download(path = sim1_output_dir)
}
```

read the file from a local folder

```{r load-sim1, eval=T}
sx1 <- read_csv("Tutorial_4_planning/data/sim1/sim1_data.csv") %>%
  mutate(exp=factor(exp),
         subj_n=factor(subj_n),
         rep_n=factor(rep_n),
         b6c = factor(b6c,
                      levels = c("-1.94", "-1.5", "-0.67"),
                      labels = c("25%", "50%", "75%")),
         subj=factor(subj),
         bin=factor(bin))
sx1
```

at the exp level

```{r exp-summary-sim1, eval=T}
sx1_exp <- sx1 %>% 
  group_by(exp, subj_n, rep_n, b6c, condition, bin) %>% 
  summarise(exp_mean_outcome = mean(mean_outcome, na.rm = TRUE),
            exp_mean_sum=mean(sum, na.rm = TRUE),
            exp_sum=sum(sum),
            n = length(unique(subj)), 
            sd=sd(mean_outcome, na.rm = TRUE),
            sem = (sd/sqrt(length(unique(subj)))),
            .groups = "drop")
sx1_exp
```

at the sim level

```{r sim-summary-sim1, eval=T}
sx1_sim <- sx1_exp %>% 
  group_by(subj_n, rep_n, b6c, condition, bin) %>% 
  summarise(sim_mean_outcome = mean(exp_mean_outcome, na.rm = TRUE),
            sim_mean_sum = mean(exp_mean_sum, na.rm = TRUE),
            sim_sum = mean(exp_sum, na.rm = TRUE),
            n = length(unique(exp)), 
            sd=sd(exp_mean_outcome, na.rm = TRUE),
            sem = (sd/sqrt(length(unique(exp)))),
            .groups = "drop")
sx1_sim
```

a quick plot of the data

```{r sim1-violin, eval=T}
p9.1 <- ggplot(sx1_sim, 
                aes(x = bin, y = sim_mean_outcome, 
                    colour = condition)) +
  geom_jitter(data=sx1_exp, aes(y = exp_mean_outcome),
              alpha=0.5, width = 0.1, height = 0) +
  geom_line(aes(group = condition)) + 
  geom_point(size=2, colour = "black") +
  geom_errorbar(aes(ymin = sim_mean_outcome-sem*1.96, ymax = sim_mean_outcome+sem*1.96),
                width=.2, colour = "black") +
  scale_colour_brewer(palette = "Dark2") +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
  labs(x="time bin",
       y="outcome") + 
  facet_grid(fct_rev(rep_n)~b6c)
p9.1

ggsave ("Tutorial_4_planning/figures/sim1/sim_violin.jpeg",
        width = 10, height = 8, dpi=800)
```

ok, look at some values in bin 6

at the exp level

```{r sim1-bin6-exp-values}
sx1_exp_check <- sx1_exp %>%
  filter(bin %in% c(6)) %>% 
  select(exp, rep_n, b6c, condition, bin, exp_mean_outcome) %>%
  pivot_wider(names_from = "condition",
              values_from = "exp_mean_outcome") %>% 
  group_by(exp, rep_n, b6c, bin) %>% 
  mutate(ratio = cond2/cond1) 
head(sx1_exp_check)
```

and at the sim level

```{r sim1-bin6-sim-values}
sx1_sim_check <- sx1_exp_check %>%
  group_by(rep_n, b6c, bin) %>% 
  summarise(sim_ratio = mean(ratio),
            sim_sd=sd(ratio),
            sim_sem=sim_sd/sqrt(1000),
            .groups = "drop")
head(sx1_sim_check)
```

plot ratio values

```{r plot-ratio-values}
p9.2 <- ggplot(sx1_sim_check, 
                aes(x = b6c, y = sim_ratio)) +
  geom_point(size=1.5) +
  geom_errorbar(aes(ymin = sim_ratio-sim_sem*1.96, ymax = sim_ratio+sim_sem*1.96)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), colour = "red") +
  scale_colour_brewer(palette = "Dark2") +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
  labs(x="time bin",
       y="outcome") + 
  facet_wrap(~rep_n)
p9.2
```

# 10. calculate power / precision #

basic idea - calculate the difference score in hazard between con and inc per bin and then report how many lower bound 95% quantile intervals exclude zero? This of course is not a model-based estimate. But we think it should give a reasonable guide to what we might expect.

## calculate difference scores ##

at the exp level

```{r sim1-calculate-diff-scores}
sx1_diff_exp <- sx1 %>%
  pivot_wider(id_cols = c(exp, subj_n, rep_n, b6c, subj, bin),
              names_from = "condition",
              values_from = "mean_outcome") %>% 
  mutate(diff = cond1 - cond2) %>%
  group_by(exp, subj_n, rep_n, b6c, bin) %>% 
  summarise(exp_mean_diff = mean(diff, na.rm = TRUE),
            exp_sd = sd(diff, na.rm = TRUE),
            n = n(), # n here is the total trials per condition per pid
            exp_sem = (exp_sd/sqrt(n)),
            exp_ci = 1.96*exp_sem,
            exp_dz = exp_mean_diff/exp_sd,
            .groups = "drop") 
head(sx1_diff_exp)
tail(sx1_diff_exp)
```

at the group/sim level

```{r sim1-summary-diff-scores}
sx1_diff_sim <- sx1_diff_exp %>% 
  group_by(subj_n, rep_n, b6c, bin) %>% 
  summarise(mean_diff = mean(exp_mean_diff, na.rm = TRUE),
            sd = sd(exp_mean_diff, na.rm = TRUE),
            n=n(),
            sem = (sd/sqrt((n))),
            ci = 1.96*sem) 
sx1_diff_sim
```

## plot ##

violin

diff in original units

```{r sim1-plot-sim-diffs}
p10.1 <- ggplot(sx1_diff_exp, aes(x=bin, y = exp_mean_diff,
                                colour = bin, fill = bin)) +
   geom_jitter(alpha = 0.5, width = 0.1) +
   geom_violin(alpha = 0.7) +
   geom_point(data = sx1_diff_sim, 
             aes(y = mean_diff), size = 3, position=pd2, colour="black") +
   geom_errorbar(data = sx1_diff_sim,
                aes(y = mean_diff, ymin = mean_diff-ci, ymax = mean_diff+ci),
                width=.2, position=pd2, colour = "black") +
   geom_hline(yintercept = 0, colour = "black", linetype = "dashed") +
   scale_fill_brewer(palette = "Dark2") +
   scale_colour_brewer(palette = "Dark2") +
   theme(legend.position = "none") +
   ylab("mean outcome") +
   ggtitle("difference score (cond1-cond2)") +
   facet_grid(fct_rev(rep_n)~b6c)
p10.1

ggsave("Tutorial_4_planning/figures/sim1/sim_diffs.jpeg",
       width = 10, height = 8, dpi=800)
```

plot each exp's difference score and associated 95% interval

create some factors and new variables

```{r sim1-wrangle-diffs}
sx1_diff_exp_2 <- sx1_diff_exp %>%
  group_by(exp, subj_n, rep_n, b6c, bin) %>% 
  mutate(lower = exp_mean_diff-exp_ci,
         upper = exp_mean_diff+exp_ci,
         above_zero = if_else(lower > 0, "yes", "no"), 
         above_zero = factor(above_zero, levels = c("no", "yes")))
sx1_diff_exp_2
```

calculate "power" in a quick and dirty way based on 95% CI

```{r sim1-calculate-power}
power_x1 <- sx1_diff_exp_2 %>%
  group_by(subj_n, rep_n, b6c, bin) %>%
  mutate(check = ifelse(lower > 0, 1, 0)) %>%
  summarise(power = mean(check, na.rm = TRUE)) %>% 
  filter(bin %in% c(6))
power_x1
```

plot power

```{r sim1-power-tile}
p10.2 <- ggplot(power_x1, aes(x = b6c, y=rep_n, fill = power)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.3f", power)), color = "white", size = 10) +
  scale_fill_viridis_c(limits = c(0, 1)) +
  facet_wrap(~bin) +
  labs(y="rep_n", x="b6c")
p10.2
#
ggsave ("Tutorial_4_planning/figures/sim1/power.jpeg",
        width = 10, height = 6, dpi=800)
```

join power to the df

```{r sim1-join-dfs}
sx1_diff_power <- sx1_diff_exp_2 %>%
  filter(bin %in% c(6)) %>% 
  inner_join(power_x1, by = c("subj_n", "rep_n", "b6c", "bin")) %>% 
  mutate(power = round(power * 100, 2)) 
head(sx1_diff_power)
```

plot

```{r sim1-plot-pointrange}
p10.3 <- sx1_diff_power %>%
  ggplot(aes(x = exp, y = exp_mean_diff, ymin = lower, ymax = upper)) +
  geom_pointrange(fatten = 1/2, aes(colour=above_zero)) +
  geom_hline(yintercept = 0, colour = "red") +
  scale_colour_manual(values=c("darkgrey","black")) +
  geom_text(aes(x=700, y=-0.35, label = sprintf("%.1f%s", power, "% power")), 
            color = "darkgrey", size = 5) +
  theme(legend.position = "none") +
  labs(x = "sim # (i.e., simulation index)",
       y = "hazard difference") +
  scale_x_discrete(breaks = c(250,500,750,1000)) +
  facet_grid(fct_rev(rep_n)~b6c)
p10.3

ggsave ("Tutorial_4_planning/figures/sim1/bin6_diffs.jpeg",
        width = 10, height = 8, dpi=800)
```

plot power as a bar plot

```{r sim1-power-col}
p10.4 <- ggplot(power_x1, aes(x=rep_n, y=power,
                           colour = b6c, fill = b6c)) +
  geom_col(alpha = 0.5) +
  geom_hline(yintercept = 0.8, colour = "red", linetype = "dashed") +
  geom_hline(yintercept = 0.9, colour = "black", linetype = "dashed") +
  geom_hline(yintercept = 0.95, colour = "blue", linetype = "dashed") +
  scale_colour_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  labs(title = "Simulation 1") +
  theme(legend.position = "none") +
  facet_wrap(~b6c, nrow=1)
p10.4

ggsave ("Tutorial_4_planning/figures/sim1/bin6_power_col.jpeg",
        width = 8, height = 6, dpi=800)
```

ok, on the basis of this analysis, we might make the following conclusions:

- effects of con vs inc that are as large as previously reported by Panis & Schimdt (2016) would be easy to detect with 100 trials and 10 pids. And probably less data than that also.

- Across all effect sizes, there is no benefit of adding more than 200 trials per condition.

- Within the context of these absolute hazard values, somewhere between a 50% reduction and a 75% reduction would give you 80% power to detect an effect, should it exist. e.g., a 75% reduction is nearly 100% power. A 50% reduction is 70% power. So maybe a 60% reduction would give 80% power, with N=10 and trial =200. We could run more sims to figure that out. 

- Other analyses to run - how many individual pids per exp show a clear effect of con vs inc in bin 6? This would be useful to know if we used the small-N design logic and the individual as the unit of replication.

# 11. Run a follow-up simulation #

Let's try a 50%, 60% and 70% reduction across N=10, 15 and 20.

let's check effect sizes quickly.

```{r sim2-bin6-effect-sizes}
bin6_v2 <- tibble(
  effect_id = 1:3,
  inc = rep(0.1, times=3),
  ratio = c(0.3, 0.4, 0.5),
  con = inc/ratio,
  inc_clog = clogloglink(inc),
  con_clog = clogloglink(con)
)
bin6_v2

# effect_id   inc ratio   con inc_clog con_clog
#       <int> <dbl> <dbl> <dbl>    <dbl>    <dbl>
# 1         1   0.1   0.3 0.333    -2.25   -0.903
# 2         2   0.1   0.4 0.25     -2.25   -1.25 
# 3         3   0.1   0.5 0.2      -2.25   -1.50 
```

## run the sim ##

vary effect size and pid.
Again, as before, this will take some time, probably hours...go get another beer/coffee.
And again, eval is set to FALSE, to save time for the tutorial.

```{r run-sim2, eval=FALSE}
plan(multicore)
x2 <- crossing(
  exp = 1:1000, # number of experiment replicates
  subj_n = c(10,15,20), # range of subject N
  rep_n = 200,
  b6i = -2.25,
  b6c = c(-0.90, -1.25, -1.50)
) %>%
  mutate(d = pmap(., index_sim)) %>% 
  mutate(s = map(d, index_summary)) %>% 
  select(-d)
```

## summarise the simulated data ##

```{r summary-sim2, eval=FALSE}
sx2 <- x2 %>%
  unnest(s) %>%
  mutate(exp=factor(exp),
         subj_n=factor(subj_n),
         rep_n=factor(rep_n),
         b6c = factor(b6c,
                      levels = c("-1.5", "-1.25", "-0.9"),
                      labels = c("50%", "60%", "70%")),
         subj=factor(subj),
         bin=factor(bin))
sx2
```

take a look

```{r sim2-quick-look, eval=FALSE}
head(sx2)
tail(sx2)

sx2 %>% 
  distinct(b6c)
```

save out a file 

```{r save-sim2, eval=FALSE}
## save the first simulation
write_csv(sx2, "Tutorial_4_planning/data/sim2/sim2_data.csv")
```

## calculate summary data ##

read in the file, if it is already computed.

If necessary, read it from the OSF to start with (because it is a large file). 
This step is only required to be done once.
After that, the file will be in your local directory and you can skip this chunk in the future.
To do so, set eval=TRUE in the chunk below, then re-set to FALSE once the data have been downloaded.

```{r load-osf-sim2, eval=T}
## create a relevant folder, if it does not already exist
sim2_output_dir <- here("Tutorial_4_planning/data/sim2/")

if (!dir.exists(here(sim2_output_dir))) {
  dir.create(here(sim2_output_dir))
}

## read in the data from the OSF and store it in a relevant folder, if it does not already exist
sim2_file_name <- here("Tutorial_4_planning/data/sim2/sim2_data.csv")

if (!file.exists(here(sim2_file_name))) {
  osf_retrieve_node("3dbcs") %>% 
  osf_ls_files(pattern = "Tutorial_4_planning") %>%
  osf_ls_files(pattern = "data") %>%
  osf_ls_files(pattern = "sim2") %>%
  osf_ls_files(pattern = "sim2_data") %>%
  osf_download(path = sim2_output_dir)
}
```

read the file from a local folder

```{r load-sim2, eval=T}
sx2 <- read_csv("Tutorial_4_planning/data/sim2/sim2_data.csv") %>%
  mutate(exp=factor(exp),
         subj_n=factor(subj_n),
         rep_n=factor(rep_n),
         b6c = factor(b6c,
                      levels = c("-1.5", "-1.25", "-0.9"),
                      labels = c("50%", "60%", "70%")),
         subj=factor(subj),
         bin=factor(bin))
sx2
```

at the exp level

```{r sim2-exp-summary}
sx2_exp <- sx2 %>% 
  group_by(exp, subj_n, rep_n, b6c, condition, bin) %>% 
  summarise(exp_mean_outcome = mean(mean_outcome, na.rm = TRUE),
            exp_mean_sum=mean(sum, na.rm = TRUE),
            exp_sum=sum(sum),
            n = length(unique(subj)), 
            sd=sd(mean_outcome, na.rm = TRUE),
            sem = (sd/sqrt(length(unique(subj)))),
            .groups = "drop")
sx2_exp
```

at the sim level

```{r sim2-sim-summary}
sx2_sim <- sx2_exp %>% 
  group_by(subj_n, rep_n, b6c, condition, bin) %>% 
  summarise(sim_mean_outcome = mean(exp_mean_outcome, na.rm = TRUE),
            sim_mean_sum = mean(exp_mean_sum, na.rm = TRUE),
            sim_sum = mean(exp_sum, na.rm = TRUE),
            n = length(unique(exp)), 
            sd=sd(exp_mean_outcome, na.rm = TRUE),
            sem = (sd/sqrt(length(unique(exp)))),
            .groups = "drop")
sx2_sim
```

a quick plot of the data

```{r sim2-violin}
p11.1 <- ggplot(sx2_sim, 
                aes(x = bin, y = sim_mean_outcome, 
                    colour = condition)) +
  geom_jitter(data=sx2_exp, aes(y = exp_mean_outcome),
              alpha=0.5, width = 0.1, height = 0) +
  geom_line(aes(group = condition)) + 
  geom_point(size=2, colour = "black") +
  geom_errorbar(aes(ymin = sim_mean_outcome-sem*1.96, ymax = sim_mean_outcome+sem*1.96),
                width=.2, colour = "black") +
  scale_colour_brewer(palette = "Dark2") +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
  labs(x="time bin",
       y="outcome") + 
  facet_grid(fct_rev(subj_n)~b6c)
p11.1

ggsave ("Tutorial_4_planning/figures/sim2/sim_violin.jpeg",
        width = 10, height = 8, dpi=800)
```

ok, look at some values in bin 6

at the exp level

```{r sim2-bin6-exp-values}
sx2_exp_check <- sx2_exp %>%
  filter(bin %in% c(6)) %>% 
  select(exp, subj_n, b6c, condition, bin, exp_mean_outcome) %>%
  pivot_wider(names_from = "condition",
              values_from = "exp_mean_outcome") %>% 
  group_by(exp, subj_n, b6c, bin) %>% 
  mutate(ratio = cond2/cond1) 
head(sx2_exp_check)
```

and at the sim level

```{r sim2-bin6-sim-values}
sx2_sim_check <- sx2_exp_check %>%
  group_by(subj_n, b6c, bin) %>% 
  summarise(sim_ratio = mean(ratio),
            sim_sd=sd(ratio),
            sim_sem=sim_sd/sqrt(1000),
            .groups = "drop")
head(sx2_sim_check)
```

plot ratio values

```{r sim2-ratio-values}
p11.2 <- ggplot(sx2_sim_check, 
                aes(x = b6c, y = sim_ratio)) +
  geom_point(size=1.5) +
  geom_errorbar(aes(ymin = sim_ratio-sim_sem*1.96, ymax = sim_ratio+sim_sem*1.96)) + 
  geom_hline(yintercept = c(0.3, 0.4, 0.5), colour = "red") +
  scale_colour_brewer(palette = "Dark2") +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.25)) +
  labs(x="time bin",
       y="outcome") + 
  facet_wrap(~subj_n)
p11.2
```

## calculate power / precision ##

Same basic idea as in Simulation 1.

## calculate difference scores ##

at the exp level

```{r sim2-calculate-diff-scores}
sx2_diff_exp <- sx2 %>%
  pivot_wider(id_cols = c(exp, subj_n, rep_n, b6c, subj, bin),
              names_from = "condition",
              values_from = "mean_outcome") %>% 
  mutate(diff = cond1 - cond2) %>%
  group_by(exp, subj_n, rep_n, b6c, bin) %>% 
  summarise(exp_mean_diff = mean(diff, na.rm = TRUE),
            exp_sd = sd(diff, na.rm = TRUE),
            n = n(), # n here is the total trials per condition per pid
            exp_sem = (exp_sd/sqrt(n)),
            exp_ci = 1.96*exp_sem,
            exp_dz = exp_mean_diff/exp_sd,
            .groups = "drop") 
head(sx2_diff_exp)
tail(sx2_diff_exp)
```

at the group/sim level

```{r sim2-summary-diff-scores}
sx2_diff_sim <- sx2_diff_exp %>% 
  group_by(subj_n, rep_n, b6c, bin) %>% 
  summarise(mean_diff = mean(exp_mean_diff, na.rm = TRUE),
            sd = sd(exp_mean_diff, na.rm = TRUE),
            n=n(),
            sem = (sd/sqrt((n))),
            ci = 1.96*sem) 
sx2_diff_sim
```

## plot ##

violin

diff in original units

```{r sim2-plot-sim-diffs}
p11.3 <- ggplot(sx2_diff_exp, aes(x=bin, y = exp_mean_diff,
                                colour = bin, fill = bin)) +
   geom_jitter(alpha = 0.5, width = 0.1) +
   geom_violin(alpha = 0.7) +
   geom_point(data = sx2_diff_sim, 
             aes(y = mean_diff), size = 3, position=pd2, colour="black") +
   geom_errorbar(data = sx2_diff_sim,
                aes(y = mean_diff, ymin = mean_diff-ci, ymax = mean_diff+ci),
                width=.2, position=pd2, colour = "black") +
   geom_hline(yintercept = 0, colour = "black", linetype = "dashed") +
   scale_fill_brewer(palette = "Dark2") +
   scale_colour_brewer(palette = "Dark2") +
   theme(legend.position = "none") +
   ylab("mean outcome") +
   ggtitle("difference score (cond1-cond2)") +
   facet_grid(fct_rev(subj_n)~b6c)
p11.3

ggsave("Tutorial_4_planning/figures/sim2/sim_diffs.jpeg",
       width = 10, height = 8, dpi=800)
```

plot each exp's difference score and associated 95% interval

create some factors and new variables

```{r sim2-wrangle-diffs}
sx2_diff_exp_2 <- sx2_diff_exp %>%
  group_by(exp, subj_n, rep_n, b6c, bin) %>% 
  mutate(lower = exp_mean_diff-exp_ci,
         upper = exp_mean_diff+exp_ci,
         above_zero = if_else(lower > 0, "yes", "no"), 
         above_zero = factor(above_zero, levels = c("no", "yes")))
sx2_diff_exp_2
```

calculate "power" in a quick and dirty way based on 95% CI

```{r sim2-calculate-power}
power_x2 <- sx2_diff_exp_2 %>%
  group_by(subj_n, rep_n, b6c, bin) %>%
  mutate(check = ifelse(lower > 0, 1, 0)) %>%
  summarise(power = mean(check, na.rm = TRUE)) %>% 
  filter(bin %in% c(6))
power_x2
```

plot power

```{r sim2-power-tile}
p11.4 <- ggplot(power_x2, aes(x = b6c, y=subj_n, fill = power)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.3f", power)), color = "white", size = 10) +
  scale_fill_viridis_c(limits = c(0, 1)) +
  facet_wrap(~bin) +
  labs(y="subj_n", x="b6c")
p11.4
#
ggsave ("Tutorial_4_planning/figures/sim2/power.jpeg",
        width = 10, height = 6, dpi=800)
```

join power to the df

```{r sim2-join-dfs}
sx2_diff_power <- sx2_diff_exp_2 %>%
  filter(bin %in% c(6)) %>% 
  inner_join(power_x2, by = c("subj_n", "rep_n", "b6c", "bin")) %>% 
  mutate(power = round(power * 100, 2)) 
head(sx2_diff_power)
```

plot

```{r sim2-plot-pointrange}
p11.5 <- sx2_diff_power %>%
  ggplot(aes(x = exp, y = exp_mean_diff, ymin = lower, ymax = upper)) +
  geom_pointrange(fatten = 1/2, aes(colour=above_zero)) +
  geom_hline(yintercept = 0, colour = "red") +
  scale_colour_manual(values=c("darkgrey","black")) +
  geom_text(aes(x=700, y=-0.35, label = sprintf("%.1f%s", power, "% power")), 
            color = "darkgrey", size = 5) +
  theme(legend.position = "none") +
  labs(x = "sim # (i.e., simulation index)",
       y = "hazard difference") +
  scale_x_discrete(breaks = c(250,500,750,1000)) +
  facet_grid(fct_rev(subj_n)~b6c)
p11.5

ggsave ("Tutorial_4_planning/figures/sim2/bin6_diffs.jpeg",
        width = 10, height = 8, dpi=800)
```

plot power as a bar plot

```{r sim2-power-col}
p11.6 <- ggplot(power_x2, aes(x=subj_n, y=power,
                           colour = b6c, fill = b6c)) +
  geom_col(alpha = 0.5) +
  geom_hline(yintercept = 0.8, colour = "red", linetype = "dashed") +
  geom_hline(yintercept = 0.9, colour = "black", linetype = "dashed") +
  geom_hline(yintercept = 0.95, colour = "blue", linetype = "dashed") +
  scale_colour_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  labs(title = "Simulation 2") +
  theme(legend.position = "none") +
  facet_wrap(~b6c, nrow=1)
p11.6

ggsave ("Tutorial_4_planning/figures/sim2/bin6_power_col.jpeg",
        width = 8, height = 6, dpi=800)
```


A summary of these power calculations might be as follows:

For a 70% reduction (0.3 hazard ratio), N=10, trial=200 per condition, would give nearly 100% power.
For a 60% reduction (0.4 hazard ratio), N=10, trial=200 per condition, would give nearly 90% power.
For a 50% reduction (0.5 hazard ratio), N=15, trial=200 per condition, would give over 80% power.

And a conclusion / judgment / decision process might be:

Well, like almost always, it depends on your objectives. Some considerations might be...

How much power or precision are you looking to obtain in this particular study?
Are you running multiple studies that have some form of replication built in?
What resources do you have at your disposal, such as time, money and personnel?
How easy or difficult is it to obtain the specific type of sample?

My thoughts for studies in my lab might be something like this...

Pick 0.4 or 0.5 as a target effect size since this is much smaller than that observed in published studies, then pick the corresponding N value (i.e., N=10 or N=15) that takes you over the 80% power mark.

**But**, and this is an important "but", do not solely rely on one study. Run a follow-up experiment that replicates and extends the initial result. By doing so, you avoid the Cult of the Isolated Single Study, and it reduces the reliance on any one type of power analysis. Instead, you are aiming for common patterns across two or more experiments, rather than trying to make the case that a single study has sufficient evidential value to hit some criterion mark.

# 12. Join some plots together #

Join the power column plots together from sim1 and sim2.

```{r sim1-sim2-power-plot}
p12.1 <- (p10.4 | p11.6) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(axes = 'collect')
p12.1

ggsave ("Tutorial_4_planning/figures/bin6_power_col.jpeg",
        width = 12, height = 7, dpi=800)
```

