---
title: "Tutorial_3a"
author: "Sven Panis"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this file we build a Frequentist hazard regression model for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent).

Content:

1. Load libraries.
2. Load and wrangle the person-trial-bin data set that we saved in Tutorial 1a.
3. Fit hazard model in the frequentist framework (model_M3_f).
4. Compare MLE estimates from model_M3_f with the posterior mean estimates from model_M3.

# 1. Load the libraries that we will be using

```{r load-pkg, results='hide'}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "lme4", "nlme", "broom.mixed")

lapply(pkg, library, character.only = TRUE)
```

# 2. Load and wrangle the person-trial-bin data set that we saved in Tutorial 1a.

```{r load-data}
ptb_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_hazard_modeling.csv")
print(ptb_data,n=30)
summary(ptb_data) # 26602 rows: 6 participants, trial, 3 conditions, 15 periods, and event indicator (0/1)
```

Prepare the data set.

```{r prepare-data}
# select analysis time range: (200,600] with 10 bins (time bin ranks 6 to 15)
ptb_data <- ptb_data %>% filter(period > 5)

# create factor condition, with "blank" as the reference level
ptb_data <- ptb_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))

# center TIME (period) on bin 9, and trial on trial 1000 and rescale.
ptb_data <- ptb_data %>% 
        mutate(period_9 = period - 9,
               trial_c = (trial - 1000)/1000)

head(ptb_data,n=17)
summary(ptb_data)
# 6 subjects
# condition: blank (6401), congruent (2642), incongruent (3797)
# period_9: -3 to 6
# trial_c: -0.999 to 0.540
```

# 3. 3. Fit hazard model in the frequentist framework (model_M3_f)

```{r get-data}
M3_data <- ptb_data %>% select(pid, period_9, trial_c, condition, event)
```

Model_M3_f contains the same effects as Model_M3 in Tutorial 2a.

```{r fit-model, eval=F}
model_M3_f <- glmer(event ~ 1 + condition*period_9 + 
                           condition*I(period_9^2) + 
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) +
                           (1 + condition*period_9 +
                           condition*I(period_9^2) +
                           condition*I(period_9^3) +
                           trial_c*period_9 + 
                           trial_c*I(period_9^2) + 
                           trial_c*I(period_9^3) | pid),
              
              # control parameters, data set, and complementary log-log link function
              control = glmerControl(optimizer = c("nlminbwrap"),
                                     optCtrl = list(maxfun=10000000)), 
              data=M3_data, 
              family=binomial(link="cloglog"))
```

Model_M3_f took about 3 hours.

```{r save-model, eval=F}
save(model_M3_f, file="Tutorial_3_Frequentist/model_M3_f.RData")
```

```{r effects-M3-f}
load("Tutorial_3_Frequentist/model_M3_f.RData")
summary(model_M3_f)
```

# 4. Compare MLE estimates from model_M3_f with the posterior mean estimates from model_M3.

```{r load-models}
model_M3 <- readRDS("Tutorial_2_Bayesian/models/model_M3.rds")
load("Tutorial_3_Frequentist/model_M3_f.RData")
```

```{r extract-estimates}
est <- tibble(brm = fixef(model_M3)[,c(1,3,4)],
              glmer = tidy(model_M3_f, effects = c("fixed"), conf.int=T, conf.level = 0.95)[,c(2,3,7,8)], 
              effectname = names(fixef(model_M3_f)),
              position = 1:16) 
```

```{r plot-estimates}
cols    <- c( "c1" = "#ff00ff", "c2" = "#3399ff" )
shapes  <- c("s1" = 16, "s2" = 17)

p1 <- ggplot(data = est, aes(x = position)) +
 geom_point(aes( y = brm[,"Estimate"],  color = "c1", shape = "s1")) +
 geom_point(aes( y = glmer$estimate, color = "c2", shape = "s2"), position = position_nudge(x = .1))  +
 geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], color = "c1"), show.legend = T) +
 geom_linerange(aes(ymin=glmer$conf.low, ymax=glmer$conf.high, color = "c2"), show.legend = T, position = position_nudge(x = .1)) +
 geom_hline(yintercept=0,colour = "grey", linetype = 2) +
 scale_x_continuous(breaks = 1:16,labels = est$effectname) +
 labs( x = "Effects in model M3", y = "cloglog-hazard" ) +
 scale_color_manual(name = "Model", 
                                breaks = c("c1", "c2"), 
                                values = cols,
                                labels = c("brm", "glmer")) +
 scale_shape_manual(name = "Model", 
                              breaks = c("s1", "s2"),
                              values = shapes,
                              labels = c("brm", "glmer")) +
   theme(panel.grid = element_blank(),
         panel.background = element_rect(fill="white",color="black"),
         legend.position = "top",
         axis.text.x = element_text(size=11,angle=90),
         axis.text.y = element_text(size=13),
         axis.title.x = element_text(size = 14),
         axis.title.y = element_text(size = 14),
         legend.text=element_text(size=13)) 
p1
```

```{r save-plot-estimates, eval=F}
ggsave("Tutorial_3_Frequentist/Comparison.png", height=8, width=8,dpi=300)
```
