---
title: "Tutorial_2b"
author: "sven panis"
date: "2024-10-18"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this Tutorial 2b we fit a Bayesian multilevel model to the conditional accuracy data.

# Load the libraries that we will be using

```{r load-pkg, results='hide'}
pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
         "brms", "tidybayes", "bayesplot", "future", "parallel")

lapply(pkg, library, character.only = TRUE)
```

# Set options. 

```{r set-options, results='hide'}
options(brms.backend = "cmdstanr",
        mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()
detectCores()
```

```{r check-info, results='hide'}
packageVersion("cmdstanr")
devtools::session_info("rstan")
```

# Load the person-trial data set that we saved in Tutorial 1a.

```{r load-data}
ca_data <- read_csv("Tutorial_1_descriptive_stats/data/inputfile_ca_modeling.csv")
print(ca_data,n=30)
summary(ca_data) # 2683 rows: 6 participants, trial, 3 conditions, 15 periods, and acc
```

# Prepare the data set.

```{r prepare-data}
# select analysis time range: (200,600]
ca_data <- ca_data %>% filter(drt > 5)

# create factor condition, with "blank" as the reference level
ca_data <- ca_data %>% mutate(condition = factor(condition, labels = c("blank", "congruent","incongruent")))

# center TIME (period) on bin 9, and trial on trial 1000 and rescale.
ca_data <- ca_data %>% 
        mutate(period_9 = drt - 9,
               trial_c = (trial - 1000)/1000)

head(ca_data,n=17)
summary(ca_data)
# 6 subjects
# condition: blank (1303), congruent (668), incongruent (626)
# period_9: -3 to 6
# trial_c: -0.999 to 0.540
```

# Fit model  M4_ca: third-order polynomical specification of TIME in the baseline condition (blank prime), and relax all assumptions for prime types, and trial number.

## Prepare data for M4_ca

```{r data-M4-ca}
# remove unnecessary columns before fitting a model
M4_ca_data <- ca_data %>% select(-c(bl,tr,trial,drt)) # 2597 rows
head(M4_ca_data)
```

### Set up priors for M4_ca

```{r priors-cloglog-M4-ca}
priors_M4_ca <- c(
  set_prior("normal(0,1)", class = "b"),       # weakly informative prior for beta parameters when using logit link
  set_prior("student_t(7.61, 0, 1.57)", class = "b", coef="Intercept"),      # weakly informative prior for Intercept when using logit link
  set_prior("normal(0, 1)", class = "sd"),                     # for standard deviation of random effects
  set_prior("lkj(2)", class = "cor")                           # for correlations between random effects
)
```

### Fit model M4_ca

```{r fit-model-M4-ca, eval=F}
plan(multicore)

model_M4_ca <- 
   brm(data = M4_ca_data,
       family = binomial(link="logit"),
       acc | trials(1) ~ 0 + Intercept +
                          condition*period_9*trial_c + 
                          condition*period_9*I(trial_c^2) + 
                          condition*I(period_9^2)*trial_c +
                          condition*I(period_9^2)*I(trial_c^2) +
                          condition*I(period_9^3) +
                          trial_c*I(period_9^3) +
                          (1 +  condition*period_9*trial_c +
                                condition*period_9*I(trial_c^2) + 
                                condition*I(period_9^2)*trial_c +
                                condition*I(period_9^2)*I(trial_c^2) +
                                condition*I(period_9^3) +
                                trial_c*I(period_9^3) | pid),
       prior = priors_M4_ca,
       chains = 4, cores = 4, iter = 3000, warmup = 1000, 
       control = list(adapt_delta = 0.999, step_size = 0.04, max_treedepth = 14),
       seed = 12, init = "0",
       file = "Tutorial_2_Bayesian/models/model_M4_ca")
```

Model_M4_ca took about 7 hours to run.

```{r check-model-M4-ca, eval=F}
model_M4_ca <- readRDS("Tutorial_2_Bayesian/models/model_M4_ca.rds")
summary(model_M4_ca)
```

# Display effects

```{r}
model_M4_ca <- readRDS("Tutorial_2_Bayesian/models/model_M4_ca.rds")
```

## Pareto k estimates

```{r pareto}
loo(model_M4_ca)$diagnostics %>% 
  data.frame() %>% 
  # attach the `id` values
  bind_cols(M4_ca_data) %>% 
  mutate(id = 1:n()) %>%
  
  ggplot(aes(x = id, y = pareto_k)) +
  geom_point(alpha = 3/4) + 
  geom_text(data = . %>% filter(pareto_k > .2),
            aes(x = id + 2, label = id),
            size = 3, hjust = 0) +
  theme(panel.grid = element_blank())
```

## Visualize posterior distributions of the effects of congruent and incongruent primes on logit-ca(t) relative to blank prime, for each time bin in trials 500, 1000, and 1500.

```{r post-distr}
get_variables(model_M4_ca)[1:31]

post <-as_draws_df(model_M4_ca) %>% # 8000 draws x 715 variables
   select(starts_with("b_")) %>%       # 8000 x 31
   expand_grid(period_9 = -3:6) %>%   
   mutate(period_9sq = period_9^2,
          period_9cu = period_9^3,
          trial500 = -500/1000,
          trial500sq = trial500^2,
          trial1500 = 500/1000,
          trial1500sq = trial1500^2)  

# effects for trials 500, 1000, and 1500
effects <- post %>% 
  mutate(`congruent trial 1500` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3` +
            trial1500 * `b_conditioncongruent:trial_c` + trial1500sq * `b_conditioncongruent:Itrial_cE2` +
            period_9 * trial1500 * `b_conditioncongruent:period_9:trial_c` +
            period_9sq * trial1500 * `b_conditioncongruent:trial_c:Iperiod_9E2` +
            period_9 * trial1500sq * `b_conditioncongruent:period_9:Itrial_cE2` +
            period_9sq * trial1500sq * `b_conditioncongruent:Itrial_cE2:Iperiod_9E2`, 
         
         `incongruent trial 1500` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + 
            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3` +
            trial1500 * `b_conditionincongruent:trial_c` + trial1500sq * `b_conditionincongruent:Itrial_cE2` +
            period_9 * trial1500 * `b_conditionincongruent:period_9:trial_c` +
            period_9sq * trial1500 * `b_conditionincongruent:trial_c:Iperiod_9E2` +
            period_9 * trial1500sq * `b_conditionincongruent:period_9:Itrial_cE2` +
            period_9sq * trial1500sq * `b_conditionincongruent:Itrial_cE2:Iperiod_9E2`,
    
         `congruent trial 500` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3` +
            trial500 * `b_conditioncongruent:trial_c` + trial500sq * `b_conditioncongruent:Itrial_cE2` +
            period_9 * trial500 * `b_conditioncongruent:period_9:trial_c` +
            period_9sq * trial500 * `b_conditioncongruent:trial_c:Iperiod_9E2` +
            period_9 * trial500sq * `b_conditioncongruent:period_9:Itrial_cE2` +
            period_9sq * trial500sq * `b_conditioncongruent:Itrial_cE2:Iperiod_9E2`, 
         
         `incongruent trial 500` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` +
            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3` +
            trial500 * `b_conditionincongruent:trial_c` + trial500sq * `b_conditionincongruent:Itrial_cE2` +
            period_9 * trial500 * `b_conditionincongruent:period_9:trial_c` +
            period_9sq * trial500 * `b_conditionincongruent:trial_c:Iperiod_9E2` +
            period_9 * trial500sq * `b_conditionincongruent:period_9:Itrial_cE2` +
            period_9sq * trial500sq * `b_conditionincongruent:Itrial_cE2:Iperiod_9E2`, 
    
         `congruent trial 1000` = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + 
            period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`,
          
         `incongruent trial 1000` = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` +
            period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) 
```

```{r plot-effects}
plot_facet <- effects %>%
  pivot_longer(cols=`congruent trial 1500`:`incongruent trial 1000`, names_to = "condition" ) %>%
  select(condition, period_9, value) %>%
  group_by(period_9) %>%
  
  ggplot(aes(x = period_9, y = value)) +
  stat_lineribbon(show.legend=T) +
  scale_fill_brewer() +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
                     limits = c(-3,6)) +
  scale_y_continuous(breaks = c(-40,-30,-20,-10,0,10,20,30,40), limits = c(-40,40)) +
  labs(x = "time bin", y = "logit-ca") +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=90),
        panel.background = element_rect(fill="white", color="black"),
        legend.position = "top") +
  facet_wrap(~ factor(condition,levels=c("congruent trial 500","congruent trial 1000","congruent trial 1500","incongruent trial 500","incongruent trial 1000","incongruent trial 1500")))
  
plot_facet
```

```{r save-plot-effects, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M4_ca_effects_con_incon_3trials.png", width = 8, height = 8, dpi = 300)
```

# calculate point and interval estimates, and hazard ratios

```{r table-interval-estimates}
int_c_500 <- effects %>%
  select(`congruent trial 500`, period_9) %>%
  group_by(period_9) %>%
  mean_hdi() %>%
  mutate(bin_endpoint = (period_9 + 9)*40,
        `odds ratio` = exp(`congruent trial 500`),
         mean = `congruent trial 500`,
         condition = "c500") %>%
  select(bin_endpoint,condition, mean, .lower, .upper, .width, `odds ratio`)

int_i_500 <- effects %>%
  select(`incongruent trial 500`, period_9) %>%
  group_by(period_9) %>%
  mean_hdi() %>%
  mutate(bin_endpoint = (period_9 + 9)*40,
        `odds ratio` = exp(`incongruent trial 500`),
         mean = `incongruent trial 500`,
         condition = "i500") %>%
  select(bin_endpoint, condition,mean, .lower, .upper, .width, `odds ratio`)

int_c_1000 <- effects %>%
  select(`congruent trial 1000`, period_9) %>%
  group_by(period_9) %>%
  mean_hdi() %>%
  mutate(bin_endpoint = (period_9 + 9)*40,
        `odds ratio` = exp(`congruent trial 1000`),
         mean = `congruent trial 1000`,
         condition = "c1000" ) %>%
  select(bin_endpoint,condition,mean, .lower, .upper, .width, `odds ratio`)

int_i_1000 <- effects %>%
  select(`incongruent trial 1000`, period_9) %>%
  group_by(period_9) %>%
  mean_hdi() %>%
  mutate(bin_endpoint = (period_9 + 9)*40,
        `odds ratio` = exp(`incongruent trial 1000`),
         mean = `incongruent trial 1000`,
         condition = "i1000") %>%
  select(bin_endpoint, condition,mean, .lower, .upper, .width,`odds ratio`)

int_c_1500 <- effects %>%
  select(`congruent trial 1500`, period_9) %>%
  group_by(period_9) %>%
  mean_hdi() %>%
  mutate(bin_endpoint = (period_9 + 9)*40,
        `odds ratio` = exp(`congruent trial 1500`),
         mean = `congruent trial 1500`,
         condition = "c1500" ) %>%
  select(bin_endpoint, condition, mean, .lower, .upper, .width, `odds ratio`)

int_i_1500 <- effects %>%
  select(`incongruent trial 1500`, period_9) %>%
  group_by(period_9) %>%
  mean_hdi() %>%
  mutate(bin_endpoint = (period_9 + 9)*40,
        `odds ratio` = exp(`incongruent trial 1500`),
         mean = `incongruent trial 1500`,
         condition = "i1500") %>%
  select(bin_endpoint, condition,mean, .lower, .upper, .width,  `odds ratio`)

test <- rbind(int_c_500, int_c_1000, int_c_1500,int_i_500,int_i_1000, int_i_1500)
test %>% print(n=60)
```

```{r save-table-interval-estimates, eval=F}
write_csv(test, file="Tutorial_2_Bayesian/tables/ca_effects_intervals_table.csv")
```

# plot effects on hazard ratio scale

```{r haz-ratio}
plot_facet_hr <- post %>% 
  mutate(Congruent = b_conditioncongruent + period_9 * `b_conditioncongruent:period_9` + period_9sq * `b_conditioncongruent:Iperiod_9E2` + period_9cu * `b_conditioncongruent:Iperiod_9E3`,
         Incongruent = b_conditionincongruent + period_9 * `b_conditionincongruent:period_9` + period_9sq * `b_conditionincongruent:Iperiod_9E2` + period_9cu * `b_conditionincongruent:Iperiod_9E3`) %>%
  
  pivot_longer(cols=Congruent:Incongruent, names_to = "condition" ) %>%
  select(condition, period_9, value) %>%
  mutate(value_hr = exp(value)) %>%
  group_by(period_9) %>%
  
   ggplot(aes(x = period_9, y = value_hr)) +
  stat_lineribbon(show.legend=T) +
  scale_fill_brewer() +
  geom_hline(yintercept=1, linetype="dashed", color = "red") +
  scale_x_continuous(breaks = c(-3:6), labels=c(((-3:6)+9)*40),
                     limits = c(-3,7)) +
  scale_y_continuous( limits = c(0,40)) +
  labs(x = "time bin", y = "hazard ratio") +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=90),
        panel.background = element_rect(fill="white",color="black"),
        legend.position = "top") +
  facet_wrap(~ condition)
  
plot_facet_hr

```

# Plot model-based predicted logit-ca(t) and ca(t) functions, for a single subject

```{r select-pid}
selected_subject = 6 # select pid here (1:6) manually
```

```{r help-function}
make_fitted <- function(fit, scale, ...) {
  
  fitted(fit,
         newdata = nd,
         scale = scale,
         ...) %>% 
    data.frame() %>% 
    bind_cols(nd)
}
```

# Set up new data

```{r new-data}
# define the `newdata`
nd <- tibble(pid = 1:6) %>%
  expand_grid(incon = 0:1,
                con  = 0:1,
                trial_c = c(-0.5,0,0.5),
                period_9 = -3:6) %>%
  filter(!(con ==1 & incon == 1)) %>%
  mutate(Iperiod_9E2 = period_9^2,
         Iperiod_9E3 = period_9^3,
         Itrial_cE2 = trial_c^2,
         condition = ifelse(con==1 & incon==0,"congruent",
                     ifelse(con==0 & incon==1,"incongruent", "blank")),
         condition = factor(condition, levels=c("blank" ,"congruent","incongruent"))) 
```

# First, on the logit-scale:

```{r subject-logit}
make_fitted(model_M4_ca, scale = "linear") %>% 
  mutate(cond   = rep(rep(c("N","C","I"),each=30),6),
         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("trial 500","trial 1000","trial 1500")),
         period = period_9+9) %>% 
  filter(pid==selected_subject) %>% 
  # plot
  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
             fill = condition, color = condition)) +
  geom_ribbon(alpha = 1/5, linewidth = 0) +
  geom_line() +
  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_x_continuous("time bin endpoint", breaks = 6:15, limits = c(6, 15), labels=c(6:15)*40) +
  ylab("fitted logit(ca)") +
  coord_cartesian(ylim = c(-20, 20)) +
  theme(panel.grid = element_blank(),
        panel.background = element_rect(fill="white",color="black"),
        legend.position = "top",
        axis.text.x = element_text(angle=90)) +
  facet_wrap(~ trial)
```

# Second, on ca(t) scale:

```{r subject-hazard}
make_fitted(model_M4_ca, scale = "response") %>% 
  mutate(cond   = rep(rep(c("N","C","I"),each=30),6),
         trial  = factor(trial_c,levels=c(-0.5,0.0,0.5),labels=c("trial 500","trial 1000","trial 1500")),
         period = period_9+9) %>% 
  filter(pid==selected_subject) %>% 
  # plot
  ggplot(aes(x = period, y = Estimate, ymin = Q2.5, ymax = Q97.5,
             fill = condition, color = condition)) +
  geom_ribbon(alpha = 1/5, linewidth = 0) +
  geom_line() +
  scale_fill_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_color_viridis_d(NULL, option = "A", end = .6, direction = -1) +
  scale_x_continuous("time bin endpoint", breaks = 6:15, limits = c(6, 15), labels=c(6:15)*40) +
  ylab("fitted ca(t)") +
  coord_cartesian(ylim = c(0, 1)) +
  theme(panel.grid = element_blank(),
        panel.background = element_rect(fill="white",color="black"),
        legend.position = "top",
        axis.text.x = element_text(angle=90)) +
  facet_wrap(~ trial)
```

```{r save-plot-fit-hazard, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M4_ca_effects_subject6.png", width = 10, height = 10, dpi = 300)
```



