---
title: "CreateFigure1"
author: "Rich and Sven"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file simulates the data used to create Figure 1 of the revised ms.

# Load the libraries that we will be using

## load 

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "truncnorm", "faux", "ggpmisc", "cowplot")

lapply(pkg, library, character.only = TRUE)
```

## plot settings

theme settings for ggplot

```{r plot-settings}
theme_set(
  theme_bw() +
    theme(
      text = element_text(size = 20, face = "bold"),
      title = element_text(size = 20, face = "bold"),
      legend.position = "bottom"
    )
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# Section 1: Create New Figure 1

In Figure 1 of the revision, we want to display for 1 subject the hazard and ca functions for 2 conditions, and the corresponding four means.

Steps:

## 1. Generate discrete time-to-event data for 1 subject, 2 conditions, 100 trials, and 10 time bins;

Adjust function from Rich:

```{r}
sim_haz <- function(subj_n = 10, rep_n = 200, bin_n = 10, cond_n = 2, # these can be changed when calling the f
                    hazard_cond1 = c(0, 0.02, 0.08, 0.19, 0.42, 0.52, 0.43, 0.41, 0.41, 0.39),
                    hazard_cond2 = hazard_cond1 * c(1, 1, 1, 1, .5, 0.3, 0.7, 1, 1, 1),
                    hazard = c(hazard_cond1, hazard_cond2), #
                    ... # helps the function work with pmap() below
) {
  # set up data structure
  data <- add_random(subj = subj_n, rep = rep_n) |>
    add_within("subj", condition = c("cond1", "cond2")) |>
    add_contrast("condition", "treatment", add_cols = TRUE, colnames = c("contrast")) |>
    add_within("subj", bin = 1:bin_n) |>
    mutate(
      hazf = rep(hazard, length.out = subj_n * rep_n * bin_n * cond_n),
      event = rbinom(n(), 1, hazf)
    ) |>
    group_by(subj, rep, condition) |>
    mutate(
      cumsum = cumsum(event),
      outcomel = event == 1 & cumsum(event) < 2,
      outcomen = if_else(event == 1 & cumsum(event) == 1, 1,
        if_else(cumsum(event) >= 1, NA, 0)
      )
    ) |>
    drop_na(outcomen) |>
    ungroup()
}
```

Generate data:

```{r}
## make it reproducible
set.seed(7)

n_trials <- 200
## apply function sim_haz
data_fig1_hazard <- sim_haz(subj_n = 1, rep_n = n_trials) |>
  mutate(
    subj = factor(subj),
    trial = as.integer(str_sub(rep, start = 4, end = 6)),
    acc = ifelse(bin < 5 & condition == "cond2", 0, 1),
    acc = ifelse(bin == 5 & trial < 100, 0, acc)
  ) |>
  select(-c("contrast", "cumsum", "outcomel", "outcomen", "rep"))

head(data_fig1_hazard, n = 20)
summary(data_fig1_hazard)
```

Set up life table for generated data:

```{r}
lifetable_fig1 <- data_fig1_hazard |>
  mutate(event = str_c("event", event)) |>
  group_by(condition, bin) |>
  count(event) |>
  ungroup() |>
  pivot_wider(
    names_from = event,
    values_from = n
  ) |>
  mutate(
    event0 = ifelse(is.na(event0), 0, event0), # replace NA with 0
    event1 = ifelse(is.na(event1), 0, event1),
    risk_set = event0 + event1
  ) |> # define the risk set
  mutate(hazard = (event1 / risk_set) |> round(digits = 3)) |> # calculate hazard estimate
  mutate(
    se_haz = sqrt((hazard * (1 - hazard)) / risk_set) |> round(digits = 4), # standard error for hazard
    pmass = event1 / max(risk_set),
    se_pmass = sqrt((pmass * (1 - pmass)) / max(risk_set))
  ) |>
  group_by(condition) |>
  mutate(
    survival = (cumprod(1 - hazard)) |> round(digits = 4), # calculate survival estimate
    term = (cumsum(hazard / (risk_set * (1 - hazard)))) |> round(digits = 7), # intermediate calculation
    se_surv = (survival * sqrt(term)) |> round(digits = 5)
  ) |> # Greenwood's (1926) approximation
  ungroup()
```

Calaculate conditional accuracy:

```{r}
ca <- data_fig1_hazard |>
  filter(event == 1) |>
  group_by(condition, bin) |>
  summarize(
    ca = mean(acc) |> round(digits = 2),
    n = n(),
    .groups = "drop"
  ) |>
  ungroup() |>
  mutate(se_ca = sqrt((ca * (1 - ca)) / n) |> round(digits = 3))
ca
```


## 2. Plot hazard and conditional accuracy

```{r}
p13.1 <- ggplot(lifetable_fig1, aes(
  x = bin, y = hazard,
  colour = condition
)) +
  geom_line(aes(linetype = condition),
    linewidth = 1
  ) +
  geom_point(size = 3) +
  geom_linerange(aes(ymin = hazard - se_haz, 
                     ymax = hazard + se_haz),
                 show.legend = F
  ) +
  scale_colour_brewer(palette = "Dark2") +
  scale_x_continuous(
    breaks = 1:10,
    limits = c(1, 10),
    labels = c(1:10) * 100
  ) +
  scale_y_continuous(limits = c(0, 1)) +
  xlab("time bin (upper bound in ms)") +
  ylab("hazard") +
  theme(
    axis.text.x = element_text(angle = 90),
    panel.grid.minor.x = element_blank(),
    legend.key.size = unit(1.3, "cm")
  )

p13.1
```

```{r}
p13.2 <- ggplot(ca, aes(
  x = bin, y = ca,
  colour = condition
)) +
  geom_line(aes(linetype = condition),
    linewidth = 1
  ) +
  geom_point(size = 3) +
  geom_linerange(aes(ymin = ca - se_ca, ymax = ca + se_ca),
    show.legend = F
  ) +
  scale_colour_brewer(palette = "Dark2") +
  scale_x_continuous(
    breaks = 1:10,
    limits = c(1, 10),
    labels = c(1:10) * 100
  ) +
  scale_y_continuous(limits = c(0, 1)) +
  xlab("time bin (upper bound in ms)") +
  ylab("conditional accuracy") +
  theme(
    axis.text.x = element_text(angle = 90),
    panel.grid.minor.x = element_blank(),
    legend.key.size = unit(1.3, "cm")
  )

p13.2
```

Combine plots

```{r}
p13.3 <- (p13.1 / p13.2) +
  plot_layout(
    guides = "collect",
    axes = "collect"
  )
p13.3
```

## 3. Create RT data (assuming uniform distribution within a bin) and calculate means

use a uniform distribution to provide upper and lower bounds

```{r}
rt_dat_fig1 <- data_fig1_hazard |>
  filter(event == 1) |>
  mutate(
    u = bin * 100,
    l = u - 100,
    m = (l + u) / 2
  ) |>
  group_by(subj, trial, condition, bin) |>
  mutate(rt = runif(n = 1, min = l, max = u))

head(rt_dat_fig1)
summary(rt_dat_fig1)
```

create summary RT data per condition at the pid level for plotting mean comparisons

```{r}
rt_dat_fig1_pid <- rt_dat_fig1 |>
  group_by(condition) |>
  summarise(
    mean = mean(rt),
    sd = sd(rt),
    n = n(), #
    sem = (sd / sqrt(n))
  )
rt_dat_fig1_pid
```

create summary ACC data per condition at the pid level for plotting mean comparisons

```{r}
acc_dat_fig1_pid <- rt_dat_fig1 |>
  group_by(condition) |>
  summarise(
    mean = mean(acc),
    sd = sd(acc),
    n = n(), #
    sem = (sd / sqrt(n))
  )
acc_dat_fig1_pid
```



## 4. plot rt + accuracy

```{r}
p13.4 <- ggplot(rt_dat_fig1_pid, aes(x = condition, y = mean, fill = condition, color = condition)) +
  geom_col(alpha = 0.9, width = 0.6) +
  geom_errorbar(
    aes(ymin = mean - sem, ymax = mean + sem),
    width = 0.2,
    linewidth = 0.8,
    color = "black"
  ) +
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") +
  scale_y_continuous(limits = c(0, 600), breaks = seq(0, 600, 100)) +
  ylab("mean RT (ms)")
p13.4
```


```{r}
p13.5 <- ggplot(acc_dat_fig1_pid, aes(x = condition, y = mean, fill = condition, color = condition)) +
  geom_col(alpha = 0.9, width = 0.6) +
  geom_errorbar(
    aes(ymin = mean - sem, ymax = mean + sem),
    width = 0.2,
    linewidth = 0.8,
    color = "black"
  ) +
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") +
  scale_y_continuous(limits = c(0, 1)) +
  ylab("mean accuracy (%)")
p13.5
```

plot together 

```{r}
p13.6 <- (p13.4 / p13.5) +
  plot_layout(
    guides = "collect",
    axes = "collect"
  )
p13.6
```

Plot mean and distributional results:

```{r}
p13_left <- (p13.4 / p13.5) +
  plot_layout(guides = "collect", axes = "collect")

p13_right <- (p13.1 / p13.2) +
  plot_layout(guides = "collect", axes = "collect")

# Combine the two columns side by side
p13_final <- (p13_left | p13_right) +
  plot_annotation(tag_levels = list(c("A", " ", "B", " "))) &
  theme(plot.tag = element_text(size = 22))

p13_final
```

```{r}
ggsave("manuscript/figures/Fig1_revision2.jpeg",
  width = 10, height = 10, dpi = 800
)
```
