---
title: "CreateFiguresReview"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# 1. Load the libraries that we will be using. 

```{r load-pkg, results='hide'}
pkg <- c("tidyverse", "patchwork","cowplot","ggtext","xfun")

lapply(pkg, library, character.only = TRUE)
```

Theme settings for ggplot.

```{r plot-settings, results='hide'}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 22, face = "bold"), 
          title = element_text(size = 22, face = "bold"))
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# 2. Create Figure on experimental paradigms

create detection data

```{r}
max_RT = 500
data_detection_present <- tibble(RT = c(254,NA,234,max_RT,343,283),   
                                 trial = c(6:1))
data_detection_absent <- tibble(RT = c(max_RT,NA,max_RT,max_RT,256,max_RT),   
                                 trial = c(6:1))
```

create detection plot

```{r}
lines_p <- tibble(xstart = 0,
                  xend =   data_detection_present$RT,
                  ystart = data_detection_present$trial,
                  yend =   data_detection_present$trial)

lines_a <- tibble(xstart = 0,
                  xend =   data_detection_absent$RT,
                  ystart = data_detection_absent$trial,
                  yend =   data_detection_absent$trial)

p1 <- data_detection_present %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=0) +
  geom_vline(xintercept=max_RT,linetype = 3) +
  geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_p,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,NA,21,4,21,21), size =3, fill ="black") + # 4 for x
  scale_x_continuous(limits = c(0,max_RT)) +
  scale_y_reverse(breaks = c(1:6),limits = c(7,0), labels = c("1","2","3","4","...","M"))+
  labs(x = "RT (ms)",
       y = "Trial",
       subtitle = "Target present")+
  ggtitle("Detection task - one button") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
             text = element_text(size = 18, face = "bold"), 
        title = element_text(size = 20, face = "bold"))

p2 <- data_detection_absent %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=max_RT,linetype = 3) +
  geom_vline(xintercept=0) +
    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_a,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(4,NA,4,4,21,4), size =3, fill ="black") +
  scale_x_continuous(limits = c(0,max_RT)) +
  scale_y_reverse(breaks = c(1:6),limits = c(7,0), labels = c("1","2","3","4","...","M"))+
  labs(x = "RT (ms)",
       y = "Trial",
       subtitle = "Target absent") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        text = element_text(size = 18.01, face = "bold"), 
        title = element_text(size = 20, face = "bold"))

#title="Detection task - single button"
p_detection <- (p1/p2) + 
  # plot_annotation(title = title) +
   plot_layout(axes = "collect_x")

p_detection
```

create categorization data

```{r}
max_RT = 800
data_cat_A <- tibble(RT = c(354,NA,434,max_RT,443,383),   
                                 trial = c(6:1))
data_cat_B <- tibble(RT = c(656,NA,max_RT,437,476,543),   
                                 trial = c(6:1))
```


create categorization plot

```{r}
lines_a <- tibble(xstart = 0,
                  xend =   data_cat_A$RT,
                  ystart = data_cat_A$trial,
                  yend =   data_cat_A$trial)

lines_b <- tibble(xstart = 0,
                  xend =   data_cat_B$RT,
                  ystart = data_cat_B$trial,
                  yend =   data_cat_B$trial)

p3 <- data_cat_A %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=0) +
  geom_vline(xintercept=max_RT,linetype = 3) +
    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_a,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,NA,21,4,21,21), size =3, fill =c("white",NA,"black","black","white","black")) + # 4 for x
  scale_x_continuous(limits = c(0,max_RT)) +
  scale_y_reverse(breaks = c(1:6),limits = c(7,0), labels = c("1","2","3","4","...","M"))+
  labs(x = "RT (ms)",
       y = "Trial",
       subtitle = "Condition A")+
  ggtitle("Discrimination task - two buttons") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
             text = element_text(size = 18, face = "bold"), 
        title = element_text(size = 20, face = "bold"))

p4 <- data_cat_B %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=max_RT,linetype = 3) +
  geom_vline(xintercept=0) +
    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_b,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,NA,4,21,21,21), size =3, fill =c("black",NA,"black","white","black","white")) +
  scale_x_continuous(limits = c(0,max_RT)) +
  scale_y_reverse(breaks = c(1:6),limits = c(7,0), labels = c("1","2","3","4","...","M"))+
  labs(x = "RT (ms)",
       y = "Trial",
       subtitle = "Condition B") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        text = element_text(size = 18.01, face = "bold"), 
        title = element_text(size = 20, face = "bold"))

#title="Discrimination task - two buttons"
p_discrimination <- (p3/p4) + 
 #  plot_annotation(title = title) +
   plot_layout(axes = "collect_x")

p_discrimination
```


create unstable-perception data and plot 

```{r}
max_RT = 120
data_percept_A <- tibble(RT = c(2,26,44,62,84,113,max_RT,
                                5,31,49,67,89,107,max_RT),   
                                 trial = c(1,1,1,1,1,1,1,
                                           2,2,2,2,2,2,2))
data_percept_B <- tibble(RT = c(4,36,54,75,88,112,max_RT,
                                5,31,45,68,89,114,max_RT),   
                                 trial = c(1,1,1,1,1,1,1,
                                           2,2,2,2,2,2,2))
```


create categorization plot

```{r}
lines_a <- tibble(xstart = 0,
                  xend =   max_RT,
                  ystart = data_percept_A$trial,
                  yend =   data_percept_A$trial)

lines_b <- tibble(xstart = 0,
                  xend =   max_RT,
                  ystart = data_percept_B$trial,
                  yend =   data_percept_B$trial)

p7 <- data_percept_A %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=0) +
  geom_vline(xintercept=max_RT,linetype = 3) +
#    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_a,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,21,21,21,21,21,4,21,21,21,21,21,21,4), 
             size =3, 
             fill =c("black","white","black","white","black","white","black", "white","black","white","black","white","black","black")) + # 4 for x
  scale_x_continuous(limits = c(0,max_RT), breaks=c(0,20,40,60,80,100,120)) +
  scale_y_reverse(breaks = c(1,2),limits = c(3,0), labels = c("1","2"))+
  labs(x = "RT (s)",
       y = "Trial",
       subtitle = "Condition A")+
  ggtitle("Bistable perception task - two buttons") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
             text = element_text(size = 18, face = "bold"), 
        title = element_text(size = 20, face = "bold"),
        plot.title = element_textbox_simple())

p8 <- data_percept_B %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=max_RT,linetype = 3) +
  geom_vline(xintercept=0) +
#    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_b,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,21,21,21,21,21,4,21,21,21,21,21,21,4), size =3, fill =c("white","black","white","black","white","black","black","black","white","black","white","black","white","black")) +
  scale_x_continuous(limits = c(0,max_RT),breaks=c(0,20,40,60,80,100,120)) +
  scale_y_reverse(breaks = c(1:2),limits = c(3,0), labels = c("1","2"))+
  labs(x = "RT (s)",
       y = "Trial",
       subtitle = "Condition B") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        text = element_text(size = 18.01, face = "bold"), 
        title = element_text(size = 20, face = "bold"))

#title="Discrimination task - two buttons"
p_percept <- (p7/p8) + 
 #  plot_annotation(title = title) +
   plot_layout(axes = "collect_x")

p_percept
```






create SOA data and plot

```{r}
max_RT = 1200
data_soa_A <- tibble(RT = c(654,NA,734,max_RT,243,683),   
                               trial = c(6:1))
data_soa_B <- tibble(RT = c(956,NA,max_RT,225,776,843),   
                                 trial = c(6:1))
```


create categorization plot

```{r}
lines_a <- tibble(xstart = 0,
                  xend =   data_soa_A$RT,
                  ystart = data_soa_A$trial,
                  yend =   data_soa_A$trial)

lines_b <- tibble(xstart = 0,
                  xend =   data_soa_B$RT,
                  ystart = data_soa_B$trial,
                  yend =   data_soa_B$trial)

p5 <- data_soa_A %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=0) +
  geom_vline(xintercept=max_RT,linetype = 3) +
    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
    geom_rect(mapping=aes(xmin=400, xmax=500, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_a,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,NA,21,4,21,21), size =3, fill =c("white",NA,"black","black","white","black")) + # 4 for x
  scale_x_continuous(limits = c(0,max_RT),breaks=c(0,200,400,600,800,1000,1200), labels=c(-400,-200,0,200,400,600,800)) +
  scale_y_reverse(breaks = c(1:6),limits = c(7,0), labels = c("1","2","3","4","...","M"))+
  labs(x = "RT (ms)",
       y = "Trial",
       subtitle = "Condition A")+
  ggtitle("Discrimination task with SOA manipulation - two buttons") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
             text = element_text(size = 18, face = "bold"), 
        title = element_text(size = 20, face = "bold"),
        plot.margin = unit(c(1,1,4,1),"lines"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        plot.title = element_textbox_simple()) +
  
  coord_cartesian(clip = "off")+
  annotate(geom = "text",
           x = c(0,200,400,600,800,1000,1200),
           y = 7,
           label = c(-400,-200,0,200,400,600,800),
           vjust = 2.8,
           size=5,
           fontface="bold") +
  annotate(geom = "text",
           x = c(0,200,400,600,800,1000,1200),
           y = 7,
           label = c(0,200,400,600,800,1000,1200),
           vjust = 4.2,
           size  = 5,
           fontface="bold")
#+
#  annotate(geom = "text",
#           x = 600,
#           y = 7,
#           label = "RT (ms)",
#           vjust = 4.2,
#           size=7,
#           fontface="bold")
  
p6 <- data_soa_B %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=max_RT,linetype = 3) +
  geom_vline(xintercept=0) +
    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
    geom_rect(mapping=aes(xmin=400, xmax=500, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_b,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,NA,4,21,21,21), size =3, fill =c("white",NA,"black","black","white","black")) +
  scale_x_continuous(limits = c(0,max_RT),breaks=c(0,200,400,600,800,1000,1200), labels=c(-400,-200,0,200,400,600,800)) +
  scale_y_reverse(breaks = c(1:6),limits = c(7,0), labels = c("1","2","3","4","...","M"))+
  labs(x = "RT (ms)",
       y = "Trial",
       subtitle = "Condition B") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        text = element_text(size = 18, face = "bold"), 
        title = element_text(size = 20, face = "bold"),
        plot.margin = unit(c(1,1,4,1),"lines"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  
  coord_cartesian(clip = "off")+
  annotate(geom = "text",
           x = c(0,200,400,600,800,1000,1200),
           y = 7,
           label = c(-400,-200,0,200,400,600,800),
           vjust = 2.8,
           size=5,
            fontface="bold") +
  annotate(geom = "text",
           x = c(0,200,400,600,800,1000,1200),
           y = 7,
           label = c(0,200,400,600,800,1000,1200),
           vjust = 4.2,
           size  = 5,
           fontface="bold") +
  annotate(geom = "text",
           x = 600,
           y = 7,
           label = "RT (ms)",
           vjust = 4.2,
           size=7,
           fontface="bold")

#title="Discrimination task - two buttons"
p_discriminationSOA <- (p5/p6) 
 #  plot_annotation(title = title) +
  # plot_layout(axes = "collect_x")

p_discriminationSOA
```




create single Figure

```{r}
plot_grid(p_detection, p_discrimination,p_discriminationSOA,p_percept,
  nrow = 2,
  labels = c("A","B","C","D"),
  label_size = 22#,
  #align = "v"
)
```

save Figure

```{r}
ggsave("REVIEW1/Plot_paradigms.png",dpi = 800,width = 16,height = 17)
```



