---
title: "CreateDecisionTree"
format: html
---

# Figure 3. Decision Tree for selecting a modeling approach

```{r}
library(grid)
library(tidyverse)
library(patchwork)

rightadjust = 1

# Define node positions
nodes <- data.frame(
  id = 1:5,
  x = c(3, 4.5, 1.5, 0.0, 3)+rightadjust,
  y = c(6, 4, 4, 2, 2),
  label = c(
    "Do you know the\nshape of the\nhazard function?",
    "Parametric\nSurvival Models",
    "Do you care about\nthe shape of the\nhazard function?",
    "Cox\nRegression Model",
    "Piecewise Exponential\nModel or\n Discrete-Time Model"
  ),
  type = c("question", "outcome", "question", "outcome", "outcome")
)


# Define edges
edges <- data.frame(
  from_x = c(3, 3, 1.5, 1.5)+rightadjust,
  from_y = c(5.4, 5.4, 3.4, 3.4),
  to_x = c(4.5, 1.5, 0, 3)+rightadjust,
  to_y = c(4.7, 4.7, 2.7, 2.7),
  label = c("Yes", "No", "No", "Yes"),
  label_x = c(3.8, 2.2, .7, 2.3)+rightadjust,
  label_y = c(5.1, 5.1, 3.1, 3.1)
)

# Create the plot
ggplot() +
  # Draw edges
  geom_segment(data = edges, 
               aes(x = from_x, y = from_y, xend = to_x, yend = to_y),
               arrow = arrow(length = unit(0.4, "cm"), type = "closed", angle = 15),
               linewidth = 1.2, color = "#2c3e50") +
  
  # Add edge labels
  geom_label(data = edges,
             aes(x = label_x, y = label_y, label = label),
             fill = "white", color = "#2c3e50", 
             fontface = "bold", size = 5, label.size = 0.5) +
  
  # Draw nodes
  geom_tile(data = nodes[nodes$type == "question", ],
            aes(x = x, y = y),
            width = 2, height = 1.2, 
            fill = "#3498db", color = "#2c3e50", size = 1.5) +
  
  geom_tile(data = nodes[nodes$type == "outcome", ],
            aes(x = x, y = y),
            width = 2, height = 1.2,
            fill = "#2ecc71", color = "#2c3e50", size = 1.5) +
  
  # Add node text
  geom_text(data = nodes,
            aes(x = x, y = y, label = label),
            color = "white", fontface = "bold", size = 4.5, lineheight = 0.9) +
  
  # Formatting
  coord_cartesian(xlim = c(0, 7), ylim = c(0.5, 7)) +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0),
        plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
        panel.background = element_rect(fill="white",color="white"),
        plot.background = element_rect(fill="white",color = "white")) +
  labs(title = "Event History Analysis Model Selection Decision Tree")
```

Save the tree.

```{r}
ggsave("REVIEW2/DecisionTreeMethods.png", width = 10, height = 10, dpi = 800, bg = "transparent")

```

# Figure 2. Ways to describe a RT distribution

```{r}
#| label: plot-views-on-RT
library(gamlss.dist)

set.seed(1)

# set values
ntrials = 100000
mean = 350
sd = 60
bin_width = 20

# sample RT data
rt_data = tibble(
  rt1 = rnorm(ntrials, mean, sd),
  rt2 = rexGAUS(ntrials,mu = 300, sigma = sd, nu = 100)
) %>%
  mutate(N = 1:n())

# Create histograms
hist1 <- ggplot(rt_data, aes(x=rt1)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  theme_bw()
hist2 <- ggplot(rt_data, aes(x=rt2)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  theme_bw()

# extract bin boundaries and counts
dat1 <- layer_data(hist1)
dat2 <- layer_data(hist2)

# create life tables
lt1 <- tibble(
  bin = 1:dim(dat1)[1],
  ncount = dat1$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat1$xmax
)

lt2 <- tibble(
  bin = 1:dim(dat2)[1],
  ncount = dat2$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat2$xmax
)

# check start of hazard
head(lt1)
head(lt2)

tail(lt1)
tail(lt2)

# replace first hazard NA with 0
lt1 <- lt1 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz)) %>%
  mutate(haz = ifelse(ncount == 0 & cumul > .999999, NA, haz))
lt2 <- lt2 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz)) %>%
  mutate(haz = ifelse(ncount == 0 & cumul > .999999, NA, haz))

tail(lt1,n=30)
tail(lt2)

# Define color palette
cbPalette <- c("#009E73", "#D55E00")
  
# discrete-time hazard functions
p1 <- tibble(
  bin = lt1$bin,
  normal = lt1$haz,
  exgaussian = lt2$haz
) %>% 
  pivot_longer(cols = normal:exgaussian,
               values_to  = "hazard probability"
               ) %>%
  rename(distribution = name) %>%
  # plot
  ggplot(aes(x=bin,y=`hazard probability`,group=distribution,colour=distribution)) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_color_manual(values = cbPalette) +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(0,seq(3,60,3)), 
                     limits = c(-0.4,60.1), 
                     labels = c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.text = element_text(size=12),
        legend.title = element_text(size=14))
p1

# prob. mass functions
p2 <- tibble(
  bin = lt1$bin,
  normal = lt1$pmass,
  exgaussian = lt2$pmass
) %>% 
  pivot_longer(cols = normal:exgaussian,
               values_to  = "probability mass"
               ) %>%
  rename(distribution = name) %>%
  # plot
  ggplot(aes(x=bin,y=`probability mass`,group=distribution,colour=distribution)) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_color_manual(values = cbPalette) +
  scale_y_continuous(limits = c(0,.2), 
                     breaks = c(0,.1,.2),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(0,seq(3,60,3)), 
                     limits = c(-0.4,60.1), 
                     labels = c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.x = element_blank(),
        legend.text = element_text(size=12),
        legend.title = element_text(size=14))
p2
# survivor functions
p3 <- tibble(
  bin = lt1$bin,
  normal = lt1$surv,
  exgaussian = lt2$surv
) %>% 
  pivot_longer(cols = normal:exgaussian,
               values_to  = "survival probability"
               ) %>%
  rename(distribution = name) %>%
  # plot
  ggplot(aes(x=bin,y=`survival probability`,group=distribution,colour=distribution)) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_color_manual(values = cbPalette) +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02)))+
  scale_x_continuous(breaks = c(0,seq(3,60,3)), 
                     limits = c(-0.4,60.1), 
                     labels = c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.text = element_text(size=12),
        legend.title = element_text(size=14))
p3

# cum. prob. functions
p4 <- tibble(
  bin = lt1$bin,
  normal = lt1$cumul,
  exgaussian = lt2$cumul
) %>% 
  pivot_longer(cols = normal:exgaussian,
               values_to  = "cumulative probability"
               ) %>%
  rename(distribution = name) %>%
  ggplot(aes(x=bin,y=`cumulative probability`,group=distribution,colour=distribution)) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
   scale_color_manual(values = cbPalette) +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02)))+
  scale_x_continuous(breaks = c(0,seq(3,60,3)), 
                     limits = c(-0.4,60.1), 
                     labels = c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.text = element_text(size=12),
        legend.title = element_text(size=14))
p4
             
# Combine         
p4_theme <- p4 + theme(legend.position = "none")
p2_theme <- p2 + theme(legend.position = "none")
p3_theme <- p3 + theme(legend.position = "none")

(p2_theme / p4_theme / p3_theme / p1) +
    plot_annotation(title = "Four ways to describe a distribution of event times:") +
    plot_layout(guides = "collect",
                axes = "collect_x") & 
    theme(legend.position = "bottom")
```

Save the plot

```{r}
ggsave("REVIEW2/Plot_distributions.png", width = 10, height = 10, dpi = 800)
```

# Figure 4. Members of different families

We plot various members of the exGaussian, gamma, and the Ratcliff diffusion model.

## gamma

```{r}
#| label: gamma
library(gamlss.dist)

set.seed(2)

# values
ntrials = 100000
mean = 350
sd = 60
bin_width = 20

# sample
rt_data = tibble(
  rt1 = rGA(ntrials, mu = 350, sigma = .3),  # gamma distr
  rt2 = rGA(ntrials, mu = 325, sigma = .3),
  rt3 = rGA(ntrials, mu = 350, sigma = .2),
  rt4 = rGA(ntrials, mu = 325, sigma = .2),
  rt5 = rGA(ntrials, mu = 450, sigma = .4),
  rt6 = rGA(ntrials, mu = 450, sigma = .5)
) %>%
  mutate(N = 1:n())

# histograms
hist1 <- ggplot(rt_data, aes(x=rt1)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist2 <- ggplot(rt_data, aes(x=rt2)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist3 <- ggplot(rt_data, aes(x=rt3)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist4 <- ggplot(rt_data, aes(x=rt4)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist5 <- ggplot(rt_data, aes(x=rt5)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist6 <- ggplot(rt_data, aes(x=rt6)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()

# extract bin boundaries and counts
dat1 <- layer_data(hist1)
dat2 <- layer_data(hist2)
dat3 <- layer_data(hist3)
dat4 <- layer_data(hist4)
dat5 <- layer_data(hist5)
dat6 <- layer_data(hist6)

# create life tables
lt1 <- tibble(
  bin = 1:dim(dat1)[1],
  ncount = dat1$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat1$xmax
)
lt2 <- tibble(
  bin = 1:dim(dat2)[1],
  ncount = dat2$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat2$xmax
)
lt3 <- tibble(
  bin = 1:dim(dat3)[1],
  ncount = dat3$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat3$xmax
)
lt4 <- tibble(
  bin = 1:dim(dat4)[1],
  ncount = dat4$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat4$xmax
)
lt5 <- tibble(
  bin = 1:dim(dat5)[1],
  ncount = dat5$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat5$xmax
)
lt6 <- tibble(
  bin = 1:dim(dat6)[1],
  ncount = dat6$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat6$xmax
)

# check heads
head(lt1)
head(lt2)
head(lt3)
head(lt4)
head(lt5)
head(lt6)

# check tails
tail(lt1, n=20)
tail(lt2, n=20)
tail(lt3)
tail(lt4, n=30)
tail(lt5)
tail(lt6)

# find where hazard reaches 1, and set hazard to NA for later bins
lt1 <- lt1 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt1 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt1,n=20)

lt2 <- lt2 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt2 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt2,n=20)

lt3 <- lt3 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt3 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt3,n=20)

lt4 <- lt4 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt4 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt4,n=30)

#lt5 <- lt5 %>% 
#  mutate(haz = ifelse(bin == 1, 0, haz),
#         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#         test = ifelse(bin < (lt5 %>% filter(test == 1) %>% pull(bin)), 1, test),
#         haz = ifelse(test == 0, NA, haz))
#tail(lt5,n=20)
  

#lt6 <- lt6 %>% 
#  mutate(haz = ifelse(bin == 1, 0, haz),
#         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#         test = ifelse(bin < (lt6 %>% filter(test == 1) %>% pull(bin)), 1, test),
#         haz = ifelse(test == 0, NA, haz))
#tail(lt6,n=20)

# Define color palette
cbPalette <- c("#009E73", "#D55E00")
  
p1 <- tibble(
  bin = lt1$bin,
  `mu=350, sigma=.3` = lt1$haz,
  `mu=325, sigma=.3` = lt2$haz,
  `mu=300, sigma=.3` = lt3$haz,
  `mu=350, sigma=.2` = lt4$haz,
  `mu=325, sigma=.2` = lt5$haz,
  `mu=400, sigma=.5` = lt6$haz
) %>% 
  pivot_longer(cols = `mu=350, sigma=.3`:`mu=400, sigma=.5`,
               values_to  = "hazard"
               ) %>%
  rename(parameters = name) %>%
  # plot
  ggplot(aes(x=bin,y=hazard,group=parameters,colour=parameters)) +
  geom_line() + 
  geom_point(size = .5) +
  theme_bw() +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(0,seq(3,60,3)), 
                     limits = c(-0.4,60.1), 
                     labels = c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)",
       y = "hazard probability",
       title = "gamma distribution") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = "right")
p1
```

## exGaussian

```{r}
#| label: ex-gaussian
library(gamlss.dist)

# values
ntrials = 200000
mean = 350
sd = 60
bin_width = 20

# sample
rt_data = tibble(
  rt1 = rexGAUS(ntrials, mu = 350, sigma = 60, nu = 100),
  rt2 = rexGAUS(ntrials, mu = 325, sigma = 60, nu = 50),
  rt3 = rexGAUS(ntrials, mu = 300, sigma = 55, nu = 100),
  rt4 = rexGAUS(ntrials, mu = 350, sigma = 55, nu = 50),
  rt5 = rexGAUS(ntrials, mu = 325, sigma = 50, nu = 100),
  rt6 = rexGAUS(ntrials, mu = 300, sigma = 50, nu = 50)
) %>%
  mutate(N = 1:n())

# histograms
hist1 <- ggplot(rt_data, aes(x=rt1)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist2 <- ggplot(rt_data, aes(x=rt2)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist3 <- ggplot(rt_data, aes(x=rt3)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist4 <- ggplot(rt_data, aes(x=rt4)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist5 <- ggplot(rt_data, aes(x=rt5)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist6 <- ggplot(rt_data, aes(x=rt6)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()

# extract bin boundaries and counts
dat1 <- layer_data(hist1)
dat2 <- layer_data(hist2)
dat3 <- layer_data(hist3)
dat4 <- layer_data(hist4)
dat5 <- layer_data(hist5)
dat6 <- layer_data(hist6)

# create life tables
lt1 <- tibble(
  bin = 1:dim(dat1)[1],
  ncount = dat1$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat1$xmax
)
lt2 <- tibble(
  bin = 1:dim(dat2)[1],
  ncount = dat2$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat2$xmax
)
lt3 <- tibble(
  bin = 1:dim(dat3)[1],
  ncount = dat3$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat3$xmax
)
lt4 <- tibble(
  bin = 1:dim(dat4)[1],
  ncount = dat4$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat4$xmax
)
lt5 <- tibble(
  bin = 1:dim(dat5)[1],
  ncount = dat5$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat5$xmax
)
lt6 <- tibble(
  bin = 1:dim(dat6)[1],
  ncount = dat6$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat6$xmax
)

# check heads
head(lt1)
head(lt2)
head(lt3)
head(lt4)
head(lt5)
head(lt6)

# check tails
tail(lt1, n=20)
tail(lt2, n=20)
tail(lt3)
tail(lt4, n=20)
tail(lt5)
tail(lt6, n=20)

# find where hazard reaches 1, and set hazard to NA for later bins
#lt1 <- lt1 %>% 
#  mutate(haz = ifelse(bin == 1, 0, haz),
#         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#         test = ifelse(bin < (lt1 %>% filter(test == 1) %>% pull(bin)), 1, test),
#         haz = ifelse(test == 0, NA, haz))
#tail(lt1,n=20)

lt2 <- lt2 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt2 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt2,n=20)

#lt3 <- lt3 %>% 
#  mutate(haz = ifelse(bin == 1, 0, haz),
#         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#         test = ifelse(bin < (lt3 %>% filter(test == 1) %>% pull(bin)), 1, test),
#         haz = ifelse(test == 0, NA, haz))
#tail(lt3,n=20)

lt4 <- lt4 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt4 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt4,n=30)

#lt5 <- lt5 %>% 
#  mutate(haz = ifelse(bin == 1, 0, haz),
#         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#         test = ifelse(bin < (lt5 %>% filter(test == 1) %>% pull(bin)), 1, test),
#         haz = ifelse(test == 0, NA, haz))
#tail(lt5,n=20)
  

lt6 <- lt6 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt6 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt6,n=20)

# Define color palette
cbPalette <- c("#009E73", "#D55E00")


# plot
p2 <- tibble(
  bin = lt1$bin,
  `mu=350, sigma=60, nu=100` = lt1$haz,
  `mu=325, sigma=60, nu=50` = lt2$haz,
  `mu=300, sigma=55, nu=100` = lt3$haz,
  `mu=350, sigma=55, nu=50` = lt4$haz,
  `mu=325, sigma=50, nu=100` = lt5$haz,
  `mu=300, sigma=50, nu=50` = lt6$haz
) %>% 
  pivot_longer(cols = `mu=350, sigma=60, nu=100`:`mu=300, sigma=50, nu=50`,
               values_to  = "hazard"
               ) %>%
  rename(parameters = name) %>%
  ggplot(aes(x=bin,y=hazard,group=parameters,colour=parameters)) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(0,seq(3,60,3)), 
                     limits = c(-0.4,60.1), 
                     labels = c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)",
       y = "hazard probability",
       title = "ex-gaussian distribution") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = "right")
p2
```

## diffusion

```{r}
#| label: diffusion
library(rtdists)

# values
ntrials = 100000
bin_width = 20
# a : Threshold separation. Typical range: 0.5 < a < 2
# v : drift rate. Typical range: -5 < v < 5
# t0 : non-decision time in s. Typical range: 0.1 < t0 < 0.5
# z : starting point. Default is 0.5*a (i.e., no bias)
# d = 0
# sz : inter-trial-variability of starting point.Values different from 0 can predict fast errors (but can slow computation considerably). Typical range: 0 < sz < 0.5. Default is 0.
# sv : inter-trial-variability of drift rate. Standard deviation of a normal distributionwith mean v describing the distribution of actual drift rates from specific trials.Values different from 0 can predict slow errors. Typical range: 0 < sv < 2.Default is 0.
# st0 : inter-trial var. of non-decisional component
# s = 1


# sample
rt1 = rdiffusion(ntrials, a=1.0, v=2.0, t0=0.2, z, d = 0, sz = 0, sv = 0,
               st0 = 0, s = 1, precision = 3, 
               stop_on_error = TRUE, maxt = 20, method = "fastdm")
  
rt2 = rdiffusion(ntrials, a=1.0, v=1.5, t0=0.2, z, d = 0, sz = 0, sv = 0,
               st0 = 0, s = 1, precision = 3, 
               stop_on_error = TRUE, maxt = 20, method = "fastdm")
  
rt3 = rdiffusion(ntrials, a=1.0, v=1.0, t0=0.2, z, d = 0, sz = 0, sv = 0,
               st0 = 0, s = 1, precision = 3, 
               stop_on_error = TRUE, maxt = 20, method = "fastdm")
  
rt4 = rdiffusion(ntrials, a=1.2, v=1.5, t0=0.3, z, d = 0, sz = 0, sv = 0,
               st0 = 0, s = 1, precision = 3, 
               stop_on_error = TRUE, maxt = 20, method = "fastdm")
  
rt5 = rdiffusion(ntrials, a=1.0, v=1.5, t0=.3, z, d = 0, sz = .4, sv = 0,
               st0 = 0, s = 1, precision = 3, 
               stop_on_error = TRUE, maxt = 20, method = "fastdm")
  
rt6 = rdiffusion(ntrials, a=1.5, v=1.0, t0=.3, z, d = 0, sz = .4, sv = 0,
               st0 = 0, s = 1, precision = 3, 
               stop_on_error = TRUE, maxt = 20, method = "fastdm")

rt_data <- tibble(
    rt1 = rt1$rt*1000,
    rt2 = rt2$rt*1000,
    rt3 = rt3$rt*1000,
    rt4 = rt4$rt*1000,
    rt5 = rt5$rt*1000,
    rt6 = rt6$rt*1000
  ) %>%
  mutate(N = 1:n()) 


# histograms
hist1 <- ggplot(rt_data, aes(x=rt1)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist2 <- ggplot(rt_data, aes(x=rt2)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist3 <- ggplot(rt_data, aes(x=rt3)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist4 <- ggplot(rt_data, aes(x=rt4)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist5 <- ggplot(rt_data, aes(x=rt5)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist6 <- ggplot(rt_data, aes(x=rt6)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()

# extract bin boundaries and counts
dat1 <- layer_data(hist1)
dat2 <- layer_data(hist2)
dat3 <- layer_data(hist3)
dat4 <- layer_data(hist4)
dat5 <- layer_data(hist5)
dat6 <- layer_data(hist6)

# create life tables
lt1 <- tibble(
  bin = 1:dim(dat1)[1],
  ncount = dat1$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat1$xmax
)
lt2 <- tibble(
  bin = 1:dim(dat2)[1],
  ncount = dat2$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat2$xmax
)
lt3 <- tibble(
  bin = 1:dim(dat3)[1],
  ncount = dat3$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat3$xmax
)
lt4 <- tibble(
  bin = 1:dim(dat4)[1],
  ncount = dat4$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat4$xmax
)
lt5 <- tibble(
  bin = 1:dim(dat5)[1],
  ncount = dat5$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat5$xmax
)
lt6 <- tibble(
  bin = 1:dim(dat6)[1],
  ncount = dat6$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat6$xmax
)


# plot
p3 <- tibble(
  bin = lt1$bin,
  `a=1.0,v=2.0,t0=0.2,sz=0` = lt1$haz,
  `a=1.0,v=1.5,t0=0.2,sz=0` = lt2$haz,
  `a=1.0,v=1.0,t0=0.2,sz=0` = lt3$haz,
  `a=1.2,v=1.5,t0=0.3,sz=0` = lt4$haz,
  `a=1.0,v=1.5,t0=0.3,sz=0.4` = lt5$haz,
  `a=1.5,v=1.0,t0=0.3,sz=0.4` = lt6$haz
) %>% 
  pivot_longer(cols = `a=1.0,v=2.0,t0=0.2,sz=0`:`a=1.5,v=1.0,t0=0.3,sz=0.4`,
               values_to  = "hazard"
               ) %>%
  rename(parameters = name) %>%
#  mutate(distribution = factor(distribution,
#                               levels = #c("diff_10_20","diff_10_15","diff_10_10","diff_14_15","diff_12_15","diff_08_15"))) #%>%
  
  ggplot(aes(x=bin,y=hazard,group=parameters,colour=parameters)) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_y_continuous(limits = c(0,.2), 
                     breaks = c(0, 0.05, 0.1,.15,.2),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(0,seq(3,60,3)), 
                     limits = c(-0.4,60.1), 
                     labels = c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)",
       y = "hazard probability",
       title = "Ratcliff diffusion model") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = "right")
p3
```
   


## lognormal

```{r}
#| label: log-normal
library(gamlss.dist)

set.seed(2)
# values
ntrials = 200000
bin_width = 20

# sample
rt_data = tibble(
  rt1 = rlnorm(ntrials, 5.8, 0.5),
  rt2 = rlnorm(ntrials, 5.6, 0.3),
  rt3 = rlnorm(ntrials, 5.4, 0.2),
  rt4 = rlnorm(ntrials, 5.4, 0.3),
  rt5 = rlnorm(ntrials, 5.4, 0.4),
  rt6 = rlnorm(ntrials, 5.4, 0.5)
) %>%
  mutate(N = 1:n())

# histograms
hist1 <- ggplot(rt_data, aes(x=rt1)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist2 <- ggplot(rt_data, aes(x=rt2)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist3 <- ggplot(rt_data, aes(x=rt3)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist4 <- ggplot(rt_data, aes(x=rt4)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist5 <- ggplot(rt_data, aes(x=rt5)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist6 <- ggplot(rt_data, aes(x=rt6)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()

# extract bin boundaries and counts
dat1 <- layer_data(hist1)
dat2 <- layer_data(hist2)
dat3 <- layer_data(hist3)
dat4 <- layer_data(hist4)
dat5 <- layer_data(hist5)
dat6 <- layer_data(hist6)

# create life tables
lt1 <- tibble(
  bin = 1:dim(dat1)[1],
  ncount = dat1$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat1$xmax
)
lt2 <- tibble(
  bin = 1:dim(dat2)[1],
  ncount = dat2$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat2$xmax
)
lt3 <- tibble(
  bin = 1:dim(dat3)[1],
  ncount = dat3$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat3$xmax
)
lt4 <- tibble(
  bin = 1:dim(dat4)[1],
  ncount = dat4$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat4$xmax
)
lt5 <- tibble(
  bin = 1:dim(dat5)[1],
  ncount = dat5$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat5$xmax
)
lt6 <- tibble(
  bin = 1:dim(dat6)[1],
  ncount = dat6$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat6$xmax
)

# check heads
head(lt1)
head(lt2)
head(lt3)
head(lt4)
head(lt5)
head(lt6)

# check tails
tail(lt1, n=20)
tail(lt2, n=20)
tail(lt3, n=20)
tail(lt4, n=30)
tail(lt5)
tail(lt6)

# find where hazard reaches 1, and set hazard to NA for later bins
#lt1 <- lt1 %>% 
#  mutate(haz = ifelse(bin == 1, 0, haz),
#         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#         test = ifelse(bin < (lt1 %>% filter(test == 1) %>% pull(bin)), 1, test),
#         haz = ifelse(test == 0, NA, haz))
#tail(lt1,n=20)

lt2 <- lt2 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt2 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt2,n=20)

lt3 <- lt3 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt3 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt3,n=20)

lt4 <- lt4 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt4 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt4,n=30)

lt5 <- lt5 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt5 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt5,n=20)
  

#lt6 <- lt6 %>% 
#  mutate(haz = ifelse(bin == 1, 0, haz),
#         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#         test = ifelse(bin < (lt6 %>% filter(test == 1) %>% pull(bin)), 1, test),
#         haz = ifelse(test == 0, NA, haz))
#tail(lt6,n=20)



# plot
p4 <- tibble(
  bin = lt1$bin,
  `meanlog=5.8, sdlog=0.5` = lt1$haz,
  `meanlog=5.6, sdlog=0.3` = lt2$haz,
  `meanlog=5.4, sdlog=0.2` = lt3$haz,
  `meanlog=5.4, sdlog=0.3` = lt4$haz,
  `meanlog=5.4, sdlog=0.4` = lt5$haz,
  `meanlog=5.4, sdlog=0.5` = lt6$haz
) %>% 
  pivot_longer(cols = `meanlog=5.8, sdlog=0.5`:`meanlog=5.4, sdlog=0.5`,
               values_to  = "hazard"
               ) %>%
  rename(parameters = name) %>%
  ggplot(aes(x=bin,y=hazard,group=parameters,colour=parameters)) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(0,seq(3,60,3)), 
                     limits = c(-0.4,60.1), 
                     labels = c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)",
       y = "hazard probability",
       title = "lognormal distribution") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = "right")
p4
```


## Combine plots.

```{r}
#| label: combine-plots-fig4

(p1/p2/p3/p4) 
             
p4_theme <- p4 + theme(legend.position = "none")
p2_theme <- p2 + theme(legend.position = "none")
p3_theme <- p3 + theme(legend.position = "none")

(p2_theme / p4_theme / p3_theme / p1) +
    plot_layout(guides = "collect",
                axes = "collect_x") & 
    theme(legend.position = "bottom")
```

Save the plot.

```{r}
ggsave("REVIEW2/Plot_distributions_theoretical.png", width = 10, height = 10, dpi = 800)
```

# Figure 5: Empirical hazard + ca(t) functions

Visual search: target present vs absent

```{r}

```


Spatial cueing: valid and invalid

stop-signal: no, 80, 120, 160, 200 ms






