---
title: "CreateDecisionTree"
format: html
---

In this file we create new figures for REVISION 2.

# Figure 3. Decision Tree for selecting a modeling approach

```{r}
library(grid)
library(tidyverse)
library(patchwork)

rightadjust = 1

# Define node positions
nodes <- data.frame(
  id = 1:5,
  x = c(3, 4.5, 1.5, 0.0, 3)+rightadjust,
  y = c(6, 4, 4, 2, 2),
  label = c(
    "Do you know the\nshape of the\nhazard function?",
    "Parametric\nSurvival Model",
    "Do you care about\nthe shape of the\nhazard function?",
    "Cox\nRegression Model",
    "Piecewise Exponential\nModel or\n Discrete-Time Model"
  ),
  type = c("question", "outcome", "question", "outcome", "outcome")
)


# Define edges
edges <- data.frame(
  from_x = c(3, 3, 1.5, 1.5)+rightadjust,
  from_y = c(5.4, 5.4, 3.4, 3.4),
  to_x = c(4.5, 1.5, 0, 3)+rightadjust,
  to_y = c(4.7, 4.7, 2.7, 2.7),
  label = c("Yes", "No", "No", "Yes"),
  label_x = c(3.8, 2.2, .7, 2.3)+rightadjust,
  label_y = c(5.1, 5.1, 3.1, 3.1)
)

# Create the plot
ggplot() +
  # Draw edges
  geom_segment(data = edges, 
               aes(x = from_x, y = from_y, xend = to_x, yend = to_y),
               arrow = arrow(length = unit(0.4, "cm"), type = "closed", angle = 15),
               linewidth = 1.2, color = "#2c3e50") +
  
  # Add edge labels
  geom_label(data = edges,
             aes(x = label_x, y = label_y, label = label),
             fill = "white", color = "#2c3e50", 
             fontface = "bold", size = 5, label.size = 0.5) +
  
  # Draw nodes
  geom_tile(data = nodes[nodes$type == "question", ],
            aes(x = x, y = y),
            width = 2, height = 1.2, 
            fill = "#3498db", color = "#2c3e50", size = 1.5) +
  
  geom_tile(data = nodes[nodes$type == "outcome", ],
            aes(x = x, y = y),
            width = 2, height = 1.2,
            fill = "#2ecc71", color = "#2c3e50", size = 1.5) +
  
  # Add node text
  geom_text(data = nodes,
            aes(x = x, y = y, label = label),
            color = "white", fontface = "bold", size = 4.5, lineheight = 0.9) +
  
  # Formatting
  coord_cartesian(xlim = c(0, 7), ylim = c(0.5, 7)) +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0),
        plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
        panel.background = element_rect(fill="white",color="white"),
        plot.background = element_rect(fill="white",color = "white")) +
  labs(title = "A decision tree to select among\nEvent History Analysis modeling approaches")
```

Save the tree.

```{r}
ggsave("REVIEW2/DecisionTreeMethods.png", width = 10, height = 10, dpi = 800, bg = "transparent")
```

# Figure 2. Ways to describe a RT distribution

```{r}
#| label: plot-views-on-RT
library(gamlss.dist)

set.seed(1)

# set values
ntrials = 100000
mean = 350
sd = 60
bin_width = 20

# sample RT data
rt_data = tibble(
  rt1 = rnorm(ntrials, mean, sd),
  rt2 = rexGAUS(ntrials,mu = 300, sigma = sd, nu = 100)
) %>%
  mutate(N = 1:n())

# Create histograms
hist1 <- ggplot(rt_data, aes(x=rt1)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  theme_bw()
hist2 <- ggplot(rt_data, aes(x=rt2)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  theme_bw()

# extract bin boundaries and counts
dat1 <- layer_data(hist1)
dat2 <- layer_data(hist2)

# create life tables
lt1 <- tibble(
  bin = 1:dim(dat1)[1],
  ncount = dat1$count,
  cumcount = cumsum(ncount),
  riskset2 = ntrials - lag(cumcount),
  riskset = ifelse(bin==1, ntrials, riskset2),
  pmass = ncount/ntrials,
  se_pmass = sqrt((pmass * (1 - pmass)) / max(riskset)),
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz2 = pmass/lag(surv, 1),
  haz = ifelse(bin==1, 0, haz2),
  se_haz = sqrt((haz * (1 - haz)) / riskset),
  temp_surv = cumsum(haz / (riskset * (1 - haz))),
  se_surv = surv * sqrt(temp_surv),
  endp = bin*bin_width,
  endpo = dat1$xmax
)

lt2 <- tibble(
  bin = 1:dim(dat2)[1],
  ncount = dat2$count,
  cumcount = cumsum(ncount),
  riskset2 = ntrials - lag(cumcount),
  riskset = ifelse(bin==1, ntrials, riskset2),
  pmass = ncount/ntrials,
  se_pmass = sqrt((pmass * (1 - pmass)) / max(riskset)),
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz2 = pmass/lag(surv, 1),
  haz = ifelse(bin==1, 0, haz2),
  se_haz = sqrt((haz * (1 - haz)) / riskset),
  temp_surv = cumsum(haz / (riskset * (1 - haz))),
  se_surv = surv * sqrt(temp_surv),
  endp = bin*bin_width,
  endpo = dat2$xmax
)

# check start of hazard
head(lt1)
head(lt2)

tail(lt1)
tail(lt2)

# replace first hazard NA with 0
lt1 <- lt1 %>% 
  #mutate(haz = ifelse(bin == 1, 0, haz)) %>%
  mutate(haz = ifelse(ncount == 0 & cumul > .999999, NA, haz))
#lt2 <- lt2 %>% 
  #mutate(haz = ifelse(bin == 1, 0, haz)) %>%
#  mutate(haz = ifelse(ncount == 0 & cumul > .999999, NA, haz))

tail(lt1,n=30)
tail(lt2)

# Define color palette
cbPalette <- c("#009E73", "#D55E00")

# Define x-axis labels
values <- seq(0, 1200, by = 60)
labels <- c(t(cbind(values, "", "")))
labels <- head(labels, -2)
  
# discrete-time hazard functions
p1 <- tibble(
  bin = lt1$bin,
  `normal (mean = 350, sd = 60)` = lt1$haz,
  `exgaussian (mu = 300, sigma = 60, nu = 100)` = lt2$haz,
  se_exgaus = lt2$se_haz,
  se_norm = lt1$se_haz
) %>% 
  pivot_longer(cols = `normal (mean = 350, sd = 60)`:`exgaussian (mu = 300, sigma = 60, nu = 100)`,
               values_to  = "hazard probability"
               ) %>%
  mutate(
    se = case_when(
      name == "normal (mean = 350, sd = 60)" ~ se_norm,
      name == "exgaussian (mu = 300, sigma = 60, nu = 100)" ~ se_exgaus)
  ) %>%
  rename(distribution = name) %>%
  # plot
  ggplot(aes(x=bin,y=`hazard probability`,group=distribution,colour=distribution)) +
  geom_vline(xintercept = 0,linetype = 2) +
  geom_vline(xintercept = 18,linetype = 1,color="grey",linewidth=.2) +
  geom_vline(xintercept = 21,linetype = 1,color="grey",linewidth=.2) +
  geom_vline(xintercept = 31,linetype = 1,color="grey",linewidth=.2) +
  
  geom_linerange(aes(ymin=`hazard probability`-se, ymax=`hazard probability`+se), 
                   show.legend = F, linewidth = 0.2) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_color_manual(values = cbPalette) +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(0,seq(1,60,1)), 
                     limits = c(-0.4,60.1), 
                     labels = labels,#c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.text = element_text(size=12),
        legend.title = element_text(size=14))
p1




# prob. mass functions
p2 <- tibble(
  bin = lt1$bin,
  `normal (mean = 350, sd = 60)` = lt1$pmass,
  `exgaussian (mu = 300, sigma = 60, nu = 100)` = lt2$pmass,
  se_exgaus = lt2$se_pmass,
  se_norm = lt1$se_pmass
) %>% 
  pivot_longer(cols = `normal (mean = 350, sd = 60)`:`exgaussian (mu = 300, sigma = 60, nu = 100)`,
               values_to  = "probability mass"
               ) %>%
   mutate(
    se = case_when(
      name == "normal (mean = 350, sd = 60)" ~ se_norm,
      name == "exgaussian (mu = 300, sigma = 60, nu = 100)" ~ se_exgaus)
  ) %>%
  rename(distribution = name) %>%
  # plot
  ggplot(aes(x=bin,y=`probability mass`,group=distribution,colour=distribution)) +
    geom_vline(xintercept = 0,linetype = 2) +
  geom_vline(xintercept = 18,linetype = 1,color="grey",linewidth=.2) +
  geom_vline(xintercept = 21,linetype = 1,color="grey",linewidth=.2) +
  geom_vline(xintercept = 31,linetype = 1,color="grey",linewidth=.2) +
  
    geom_linerange(aes(ymin=`probability mass`-se, ymax=`probability mass`+se), 
                   show.legend = F, linewidth = 0.2) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_color_manual(values = cbPalette) +
  scale_y_continuous(limits = c(0,.2), 
                     breaks = c(0,.1,.2),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(0,seq(1,60,1)), 
                     limits = c(-0.4,60.1), 
                     labels = labels,#c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.x = element_blank(),
             panel.grid.major.x = element_blank(),
        legend.text = element_text(size=12),
        legend.title = element_text(size=14))
p2
# survivor functions
p3 <- tibble(
  bin = lt1$bin,
  `normal (mean = 350, sd = 60)` = lt1$surv,
  `exgaussian (mu = 300, sigma = 60, nu = 100)` = lt2$surv,
  se_exgaus = lt2$se_surv,
  se_norm = lt1$se_surv
) %>%
  pivot_longer(cols = `normal (mean = 350, sd = 60)`:`exgaussian (mu = 300, sigma = 60, nu = 100)`,
               values_to  = "survival probability"
               ) %>%
   mutate(
    se = case_when(
      name == "normal (mean = 350, sd = 60)" ~ se_norm,
      name == "exgaussian (mu = 300, sigma = 60, nu = 100)" ~ se_exgaus)
  ) %>%
  rename(distribution = name) %>%
  # plot
  ggplot(aes(x=bin,y=`survival probability`,group=distribution,colour=distribution)) +
    geom_vline(xintercept = 0,linetype = 2) +
  geom_vline(xintercept = 18,linetype = 1,color="grey",linewidth=.2) +
  geom_vline(xintercept = 21,linetype = 1,color="grey",linewidth=.2) +
  geom_vline(xintercept = 31,linetype = 1,color="grey",linewidth=.2) +
  
    geom_linerange(aes(ymin=`survival probability`-se, ymax=`survival probability`+se), 
                   show.legend = F, linewidth = 0.2) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_color_manual(values = cbPalette) +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02)))+
  scale_x_continuous(breaks = c(0,seq(1,60,1)), 
                     limits = c(-0.4,60.1), 
                     labels = labels,#c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
             panel.grid.major.x = element_blank(),
        legend.text = element_text(size=12),
        legend.title = element_text(size=14))
p3

# cum. prob. functions
p4 <- tibble(
  bin = lt1$bin,
  `normal (mean = 350, sd = 60)` = lt1$cumul,
  `exgaussian (mu = 300, sigma = 60, nu = 100)` = lt2$cumul,
  se_exgaus = lt2$se_surv,
  se_norm = lt1$se_surv
) %>% 
  pivot_longer(cols = `normal (mean = 350, sd = 60)`:`exgaussian (mu = 300, sigma = 60, nu = 100)`,
               values_to  = "cumulative probability"
               ) %>%
   mutate(
    se = case_when(
      name == "normal (mean = 350, sd = 60)" ~ se_norm,
      name == "exgaussian (mu = 300, sigma = 60, nu = 100)" ~ se_exgaus)
  ) %>%
  rename(distribution = name) %>%
  ggplot(aes(x=bin,y=`cumulative probability`,group=distribution,colour=distribution)) +
    geom_vline(xintercept = 0,linetype = 2) +
  geom_vline(xintercept = 18,linetype = 1,color="grey",linewidth=.2) +
  geom_vline(xintercept = 21,linetype = 1,color="grey",linewidth=.2) +
  geom_vline(xintercept = 31,linetype = 1,color="grey",linewidth=.2) +
  
    geom_linerange(aes(ymin=`cumulative probability`-se, ymax=`cumulative probability`+se), 
                   show.legend = F, linewidth = 0.2) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
   scale_color_manual(values = cbPalette) +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02)))+
  scale_x_continuous(breaks = c(0,seq(1,60,1)), 
                     limits = c(-0.4,60.1), 
                     labels = labels,#c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
             panel.grid.major.x = element_blank(),
        legend.text = element_text(size=12),
        legend.title = element_text(size=14))
p4
             
# Combine         
p4_theme <- p4 + theme(legend.position = "none")
p2_theme <- p2 + theme(legend.position = "none")
p3_theme <- p3 + theme(legend.position = "none")

(p2_theme / p4_theme / p3_theme / p1) +
    plot_annotation(title = "Four ways to describe the distribution of a sample of (20ms-)interval-censored event times:") +
    plot_layout(guides = "collect",
                axes = "collect_x") & 
    theme(legend.position = "bottom")
```

Save the plot

```{r}
ggsave("REVIEW2/Plot_distributions_1000.png", width = 10, height = 10, dpi = 800)
```

# Figure 4. Members of different families

We plot various members of the exGaussian, gamma, and the Ratcliff diffusion model.

## gamma

```{r}
#| label: gamma
library(gamlss.dist)

set.seed(2)

# values
ntrials = 100000
mean = 350
sd = 60
bin_width = 20

# sample
rt_data = tibble(
  rt1 = rGA(ntrials, mu = 350, sigma = .3),  # gamma distr
  rt2 = rGA(ntrials, mu = 325, sigma = .3),
  rt3 = rGA(ntrials, mu = 350, sigma = .2),
  rt4 = rGA(ntrials, mu = 325, sigma = .2),
  rt5 = rGA(ntrials, mu = 450, sigma = .4),
  rt6 = rGA(ntrials, mu = 450, sigma = .5)
) %>%
  mutate(N = 1:n())

# histograms
hist1 <- ggplot(rt_data, aes(x=rt1)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist2 <- ggplot(rt_data, aes(x=rt2)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist3 <- ggplot(rt_data, aes(x=rt3)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist4 <- ggplot(rt_data, aes(x=rt4)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist5 <- ggplot(rt_data, aes(x=rt5)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist6 <- ggplot(rt_data, aes(x=rt6)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()

# extract bin boundaries and counts
dat1 <- layer_data(hist1)
dat2 <- layer_data(hist2)
dat3 <- layer_data(hist3)
dat4 <- layer_data(hist4)
dat5 <- layer_data(hist5)
dat6 <- layer_data(hist6)

# create life tables
lt1 <- tibble(
  bin = 1:dim(dat1)[1],
  ncount = dat1$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat1$xmax
)
lt2 <- tibble(
  bin = 1:dim(dat2)[1],
  ncount = dat2$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat2$xmax
)
lt3 <- tibble(
  bin = 1:dim(dat3)[1],
  ncount = dat3$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat3$xmax
)
lt4 <- tibble(
  bin = 1:dim(dat4)[1],
  ncount = dat4$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat4$xmax
)
lt5 <- tibble(
  bin = 1:dim(dat5)[1],
  ncount = dat5$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat5$xmax
)
lt6 <- tibble(
  bin = 1:dim(dat6)[1],
  ncount = dat6$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat6$xmax
)

# check heads
head(lt1)
head(lt2)
head(lt3)
head(lt4)
head(lt5)
head(lt6)

# check tails
tail(lt1, n=20)
tail(lt2, n=20)
tail(lt3)
tail(lt4, n=30)
tail(lt5)
tail(lt6)

# find where hazard reaches 1, and set hazard to NA for later bins
lt1 <- lt1 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt1 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt1,n=20)

lt2 <- lt2 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt2 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt2,n=20)

lt3 <- lt3 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt3 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt3,n=20)

lt4 <- lt4 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt4 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt4,n=30)

#lt5 <- lt5 %>% 
#  mutate(haz = ifelse(bin == 1, 0, haz),
#         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#         test = ifelse(bin < (lt5 %>% filter(test == 1) %>% pull(bin)), 1, test),
#         haz = ifelse(test == 0, NA, haz))
#tail(lt5,n=20)
  

#lt6 <- lt6 %>% 
#  mutate(haz = ifelse(bin == 1, 0, haz),
#         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#         test = ifelse(bin < (lt6 %>% filter(test == 1) %>% pull(bin)), 1, test),
#         haz = ifelse(test == 0, NA, haz))
#tail(lt6,n=20)

# Define color palette
cbPalette <- c("#009E73", "#D55E00")
  
p1 <- tibble(
  bin = lt1$bin,
  `mu=350, sigma=.3` = lt1$haz,
  `mu=325, sigma=.3` = lt2$haz,
  `mu=300, sigma=.3` = lt3$haz,
  `mu=350, sigma=.2` = lt4$haz,
  `mu=325, sigma=.2` = lt5$haz,
  `mu=400, sigma=.5` = lt6$haz
) %>% 
  pivot_longer(cols = `mu=350, sigma=.3`:`mu=400, sigma=.5`,
               values_to  = "hazard"
               ) %>%
  rename(parameters = name) %>%
  # plot
  ggplot(aes(x=bin,y=hazard,group=parameters,colour=parameters)) +
  geom_line() + 
  geom_point(size = .5) +
  theme_bw() +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(0,seq(3,60,3)), 
                     limits = c(-0.4,60.1), 
                     labels = c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)",
       y = "hazard probability",
       title = "gamma distribution") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = "right")
p1
```

## exGaussian

```{r}
#| label: ex-gaussian
library(gamlss.dist)

# values
ntrials = 200000
mean = 350
sd = 60
bin_width = 20

# sample
rt_data = tibble(
  rt1 = rexGAUS(ntrials, mu = 350, sigma = 60, nu = 100),
  rt2 = rexGAUS(ntrials, mu = 325, sigma = 60, nu = 50),
  rt3 = rexGAUS(ntrials, mu = 300, sigma = 55, nu = 100),#2
  rt4 = rexGAUS(ntrials, mu = 350, sigma = 55, nu = 50),
  rt5 = rexGAUS(ntrials, mu = 325, sigma = 50, nu = 100),#3
  rt6 = rexGAUS(ntrials, mu = 300, sigma = 50, nu = 350)  #1
) %>%
  mutate(N = 1:n())

# histograms
hist1 <- ggplot(rt_data, aes(x=rt1)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist2 <- ggplot(rt_data, aes(x=rt2)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist3 <- ggplot(rt_data, aes(x=rt3)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist4 <- ggplot(rt_data, aes(x=rt4)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist5 <- ggplot(rt_data, aes(x=rt5)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist6 <- ggplot(rt_data, aes(x=rt6)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()

# extract bin boundaries and counts
dat1 <- layer_data(hist1)
dat2 <- layer_data(hist2)
dat3 <- layer_data(hist3)
dat4 <- layer_data(hist4)
dat5 <- layer_data(hist5)
dat6 <- layer_data(hist6)

# create life tables
lt1 <- tibble(
  bin = 1:dim(dat1)[1],
  ncount = dat1$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat1$xmax
)
lt2 <- tibble(
  bin = 1:dim(dat2)[1],
  ncount = dat2$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat2$xmax
)
lt3 <- tibble(
  bin = 1:dim(dat3)[1],
  ncount = dat3$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat3$xmax
)
lt4 <- tibble(
  bin = 1:dim(dat4)[1],
  ncount = dat4$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat4$xmax
)
lt5 <- tibble(
  bin = 1:dim(dat5)[1],
  ncount = dat5$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat5$xmax
)
lt6 <- tibble(
  bin = 1:dim(dat6)[1],
  ncount = dat6$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat6$xmax
)

# check heads
head(lt1)
head(lt2)
head(lt3)
head(lt4)
head(lt5)
head(lt6)

# check tails
tail(lt1, n=20)
tail(lt2, n=20)
tail(lt3)
tail(lt4, n=20)
tail(lt5)
tail(lt6, n=20)

# find where hazard reaches 1, and set hazard to NA for later bins
#lt1 <- lt1 %>% 
#  mutate(haz = ifelse(bin == 1, 0, haz),
#         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#         test = ifelse(bin < (lt1 %>% filter(test == 1) %>% pull(bin)), 1, test),
#         haz = ifelse(test == 0, NA, haz))
#tail(lt1,n=20)

lt2 <- lt2 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt2 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt2,n=20)

#lt3 <- lt3 %>% 
#  mutate(haz = ifelse(bin == 1, 0, haz),
#         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#         test = ifelse(bin < (lt3 %>% filter(test == 1) %>% pull(bin)), 1, test),
#         haz = ifelse(test == 0, NA, haz))
#tail(lt3,n=20)

lt4 <- lt4 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt4 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt4,n=30)

#lt5 <- lt5 %>% 
#  mutate(haz = ifelse(bin == 1, 0, haz),
#         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#         test = ifelse(bin < (lt5 %>% filter(test == 1) %>% pull(bin)), 1, test),
#         haz = ifelse(test == 0, NA, haz))
#tail(lt5,n=20)
  

# lt6 <- lt6 %>% 
#   mutate(haz = ifelse(bin == 1, 0, haz),
#          test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#          test = ifelse(bin < (lt6 %>% filter(test == 1) %>% pull(bin)), 1, test),
#          haz = ifelse(test == 0, NA, haz))
# tail(lt6,n=20)

# Define color palette
cbPalette <- c("#009E73", "#D55E00")


# plot
p2 <- tibble(
  bin = lt1$bin,
  `mu=350, sigma=60, nu=100` = lt1$haz,
  `mu=325, sigma=60, nu=50` = lt2$haz,
  `mu=300, sigma=55, nu=100` = lt3$haz,
  `mu=350, sigma=55, nu=50` = lt4$haz,
  `mu=325, sigma=50, nu=100` = lt5$haz,
  `mu=300, sigma=50, nu=350` = lt6$haz
) %>% 
  pivot_longer(cols = `mu=350, sigma=60, nu=100`:`mu=300, sigma=50, nu=350`,
               values_to  = "hazard"
               ) %>%
  rename(parameters = name) %>%
  ggplot(aes(x=bin,y=hazard,group=parameters,colour=parameters)) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(0,seq(3,60,3)), 
                     limits = c(-0.4,60.1), 
                     labels = c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)",
       y = "hazard probability",
       title = "ex-gaussian distribution") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = "right")
p2
```

## diffusion

```{r}
#| label: diffusion
library(rtdists)

# values
ntrials = 100000
bin_width = 20
# a : Threshold separation. Typical range: 0.5 < a < 2
# v : drift rate. Typical range: -5 < v < 5
# t0 : non-decision time in s. Typical range: 0.1 < t0 < 0.5
# z : starting point. Default is 0.5*a (i.e., no bias)
# d = 0
# sz : inter-trial-variability of starting point.Values different from 0 can predict fast errors (but can slow computation considerably). Typical range: 0 < sz < 0.5. Default is 0.
# sv : inter-trial-variability of drift rate. Standard deviation of a normal distributionwith mean v describing the distribution of actual drift rates from specific trials.Values different from 0 can predict slow errors. Typical range: 0 < sv < 2.Default is 0.
# st0 : inter-trial var. of non-decisional component
# s = 1

a=1.0
v=2.0
t0=.2
d=0

# sample
rt1 = rdiffusion(ntrials, a=1, v=2.0, t0=0.2, z=0.5*a, d = 0, sz = 0, sv = 0,
               st0 = 0, s = 1, precision = 3, 
               stop_on_error = TRUE, maxt = 20, method = "fastdm")
  
rt2 = rdiffusion(ntrials, a=1.0, v=1.5, t0=0.2, z=0.5*a, d = 0, sz = 0, sv = 0,
               st0 = 0, s = 1, precision = 3, 
               stop_on_error = TRUE, maxt = 20, method = "fastdm")
  
rt3 = rdiffusion(ntrials, a=1.0, v=1.0, t0=0.2, z=0.5*a, d = 0, sz = 0, sv = 0,
               st0 = 0, s = 1, precision = 3, 
               stop_on_error = TRUE, maxt = 20, method = "fastdm")
  
rt4 = rdiffusion(ntrials, a=1.2, v=1.5, t0=0.3, z=0.5*a, d = 0, sz = 0, sv = 0,
               st0 = 0, s = 1, precision = 3, 
               stop_on_error = TRUE, maxt = 20, method = "fastdm")
  
rt5 = rdiffusion(ntrials, a=1.0, v=1.5, t0=.3, z=0.5*a, d = 0, sz = .4, sv = 0,
               st0 = 0, s = 1, precision = 3, 
               stop_on_error = TRUE, maxt = 20, method = "fastdm")
  
rt6 = rdiffusion(ntrials, a=1.5, v=1.0, t0=.3, z=0.5*a, d = 0, sz = .4, sv = 0,
               st0 = 0, s = 1, precision = 3, 
               stop_on_error = TRUE, maxt = 20, method = "fastdm")

rt_data <- tibble(
    rt1 = rt1$rt*1000,
    rt2 = rt2$rt*1000,
    rt3 = rt3$rt*1000,
    rt4 = rt4$rt*1000,
    rt5 = rt5$rt*1000,
    rt6 = rt6$rt*1000
  ) %>%
  mutate(N = 1:n()) 


# histograms
hist1 <- ggplot(rt_data, aes(x=rt1)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist2 <- ggplot(rt_data, aes(x=rt2)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist3 <- ggplot(rt_data, aes(x=rt3)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist4 <- ggplot(rt_data, aes(x=rt4)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist5 <- ggplot(rt_data, aes(x=rt5)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist6 <- ggplot(rt_data, aes(x=rt6)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()

# extract bin boundaries and counts
dat1 <- layer_data(hist1)
dat2 <- layer_data(hist2)
dat3 <- layer_data(hist3)
dat4 <- layer_data(hist4)
dat5 <- layer_data(hist5)
dat6 <- layer_data(hist6)

# create life tables
lt1 <- tibble(
  bin = 1:dim(dat1)[1],
  ncount = dat1$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat1$xmax
)
lt2 <- tibble(
  bin = 1:dim(dat2)[1],
  ncount = dat2$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat2$xmax
)
lt3 <- tibble(
  bin = 1:dim(dat3)[1],
  ncount = dat3$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat3$xmax
)
lt4 <- tibble(
  bin = 1:dim(dat4)[1],
  ncount = dat4$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat4$xmax
)
lt5 <- tibble(
  bin = 1:dim(dat5)[1],
  ncount = dat5$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat5$xmax
)
lt6 <- tibble(
  bin = 1:dim(dat6)[1],
  ncount = dat6$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat6$xmax
)


# plot
p3 <- tibble(
  bin = lt1$bin,
  `a=1.0,v=2.0,t0=0.2,sz=0` = lt1$haz,
  `a=1.0,v=1.5,t0=0.2,sz=0` = lt2$haz,
  `a=1.0,v=1.0,t0=0.2,sz=0` = lt3$haz,
  `a=1.2,v=1.5,t0=0.3,sz=0` = lt4$haz,
  `a=1.0,v=1.5,t0=0.3,sz=0.4` = lt5$haz,
  `a=1.5,v=1.0,t0=0.3,sz=0.4` = lt6$haz
) %>% 
  pivot_longer(cols = `a=1.0,v=2.0,t0=0.2,sz=0`:`a=1.5,v=1.0,t0=0.3,sz=0.4`,
               values_to  = "hazard"
               ) %>%
  rename(parameters = name) %>%
#  mutate(distribution = factor(distribution,
#                               levels = #c("diff_10_20","diff_10_15","diff_10_10","diff_14_15","diff_12_15","diff_08_15"))) #%>%
  
  ggplot(aes(x=bin,y=hazard,group=parameters,colour=parameters)) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_y_continuous(limits = c(0,.2), 
                     breaks = c(0, 0.05, 0.1,.15,.2),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(0,seq(3,60,3)), 
                     limits = c(-0.4,60.1), 
                     labels = c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)",
       y = "hazard probability",
       title = "Ratcliff diffusion model") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = "right")
p3
```
   
## lognormal

```{r}
#| label: log-normal
library(gamlss.dist)

set.seed(2)
# values
ntrials = 200000
bin_width = 20

# sample
rt_data = tibble(
  rt1 = rlnorm(ntrials, 6.0, 0.7),
  rt2 = rlnorm(ntrials, 5.7, 0.7),
  rt3 = rlnorm(ntrials, 5.4, 0.2),#5.4 0.2
  rt4 = rlnorm(ntrials, 5.4, 0.3),#5.4 0.3
  rt5 = rlnorm(ntrials, 5.4, 0.4),#5.4 0.4
  rt6 = rlnorm(ntrials, 5.4, 0.7) #5.4 0.6
) %>%
  mutate(N = 1:n())

# histograms
hist1 <- ggplot(rt_data, aes(x=rt1)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist2 <- ggplot(rt_data, aes(x=rt2)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist3 <- ggplot(rt_data, aes(x=rt3)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
   scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist4 <- ggplot(rt_data, aes(x=rt4)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist5 <- ggplot(rt_data, aes(x=rt5)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()
hist6 <- ggplot(rt_data, aes(x=rt6)) +
  geom_histogram(binwidth = bin_width, boundary = bin_width) +
  scale_x_continuous(limits = c(0,1200)) +
  scale_y_continuous(limits = c(0,20000)) +
  theme_bw()

# extract bin boundaries and counts
dat1 <- layer_data(hist1)
dat2 <- layer_data(hist2)
dat3 <- layer_data(hist3)
dat4 <- layer_data(hist4)
dat5 <- layer_data(hist5)
dat6 <- layer_data(hist6)

# create life tables
lt1 <- tibble(
  bin = 1:dim(dat1)[1],
  ncount = dat1$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat1$xmax
)
lt2 <- tibble(
  bin = 1:dim(dat2)[1],
  ncount = dat2$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat2$xmax
)
lt3 <- tibble(
  bin = 1:dim(dat3)[1],
  ncount = dat3$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat3$xmax
)
lt4 <- tibble(
  bin = 1:dim(dat4)[1],
  ncount = dat4$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat4$xmax
)
lt5 <- tibble(
  bin = 1:dim(dat5)[1],
  ncount = dat5$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat5$xmax
)
lt6 <- tibble(
  bin = 1:dim(dat6)[1],
  ncount = dat6$count,
  pmass = ncount/ntrials,
  cumul = cumsum(pmass),
  surv = 1 - cumul,
  haz = pmass/lag(surv, 1),
  endp = bin*bin_width,
  endpo = dat6$xmax
)

# check heads
head(lt1)
head(lt2)
head(lt3)
head(lt4)
head(lt5)
head(lt6)

# check tails
tail(lt1, n=20)
tail(lt2, n=20)
tail(lt3, n=20)
tail(lt4, n=30)
tail(lt5)
tail(lt6)

# find where hazard reaches 1, and set hazard to NA for later bins
#lt1 <- lt1 %>% 
#  mutate(haz = ifelse(bin == 1, 0, haz),
#         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#         test = ifelse(bin < (lt1 %>% filter(test == 1) %>% pull(bin)), 1, test),
#         haz = ifelse(test == 0, NA, haz))
#tail(lt1,n=20)

# lt2 <- lt2 %>% 
#   mutate(haz = ifelse(bin == 1, 0, haz),
#          test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#          test = ifelse(bin < (lt2 %>% filter(test == 1) %>% pull(bin)), 1, test),
#          haz = ifelse(test == 0, NA, haz))
# tail(lt2,n=20)

lt3 <- lt3 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt3 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt3,n=20)

lt4 <- lt4 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt4 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt4,n=30)

lt5 <- lt5 %>% 
  mutate(haz = ifelse(bin == 1, 0, haz),
         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
         test = ifelse(bin < (lt5 %>% filter(test == 1) %>% pull(bin)), 1, test),
         haz = ifelse(test == 0, NA, haz))
tail(lt5,n=20)
  

#lt6 <- lt6 %>% 
#  mutate(haz = ifelse(bin == 1, 0, haz),
#         test = ifelse(abs(haz - 1.0) < .Machine$double.eps^0.5, 1, 0),
#         test = ifelse(bin < (lt6 %>% filter(test == 1) %>% pull(bin)), 1, test),
#         haz = ifelse(test == 0, NA, haz))
#tail(lt6,n=20)


# plot
p4 <- tibble(
  bin = lt1$bin,
  `meanlog=6.0, sdlog=0.7` = lt1$haz,
  `meanlog=5.7, sdlog=0.7` = lt2$haz,
  `meanlog=5.4, sdlog=0.2` = lt3$haz,
  `meanlog=5.4, sdlog=0.3` = lt4$haz,
  `meanlog=5.4, sdlog=0.4` = lt5$haz,
  `meanlog=5.4, sdlog=0.7` = lt6$haz
) %>% 
  pivot_longer(cols = `meanlog=6.0, sdlog=0.7`:`meanlog=5.4, sdlog=0.7`,
               values_to  = "hazard"
               ) %>%
  rename(parameters = name) %>%
  ggplot(aes(x=bin,y=hazard,group=parameters,colour=parameters)) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(0,seq(3,60,3)), 
                     limits = c(-0.4,60.1), 
                     labels = c(0,seq(3,60,3)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)",
       y = "hazard probability",
       title = "lognormal distribution") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = "right")
p4
```

## Combine plots.

```{r}
#| label: combine-plots-fig4

(p1/p2/p3/p4) 
             
p4_theme <- p4 + theme(legend.position = "none")
p2_theme <- p2 + theme(legend.position = "none")
p3_theme <- p3 + theme(legend.position = "none")

(p2_theme / p4_theme / p3_theme / p1) +
    plot_layout(guides = "collect",
                axes = "collect_x") & 
    theme(legend.position = "right")
```

Save the plot.

```{r}
ggsave("REVIEW2/Plot_distributions_theoretical.png", width = 10, height = 10, dpi = 800)
```




# Figure 5: Empirical hazard + ca(t) functions

## Visual search: target present vs absent

```{r}
pkg <- c("tidyverse", "patchwork")

lapply(pkg, library, character.only = TRUE)
```

Theme settings for ggplot.

```{r plot-settings, results='hide'}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 22, face = "bold"), 
          title = element_text(size = 22, face = "bold"))
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

Read and wrangle data.

```{r read-raw-data-vs}
raw <- read_delim(file="REVIEW2/visualsearch/ConjunctionSearchDataSetOK.txt", 
                  delim="\t",
                  col_names=T)
head(raw)
```


```{r wrangle-data-vs}
# recode accuracy
data_wr <- raw %>%
  rename(pid = Subject, rt = RT) %>%
  mutate(acc = 1 - Error) %>%
  select(pid, setsize, Targ, rt, acc) 
  
head(data_wr)
summary(data_wr)

# check number of trials
data_wr %>% group_by(pid, Targ, setsize) %>%
  summarise(N = n())

# add trial nr and factors for target-presence and set-size
data_wr <- data_wr %>%
  group_by(pid) %>%
  mutate(trial = 1:n(),
         setsize = factor(setsize,
                          levels = c(3,6,12,18),
                          labels = c("3","6","12","18")),
         targetpresence = factor(Targ, levels = c(1,0),
                         labels = c("Target present","Target absent"))) %>%
  mutate(condition1 = setsize,
         condition2 = targetpresence) %>%
  select(pid, trial, condition1, condition2, rt, acc)

summary(data_wr)
head(data_wr)
```

set up life tables 

```{r load-functions-vs}
source("REVIEW2/visualsearch/custom_functions_visualsearch.R")
```

Apply the custom functions

```{r apply-functions-vs}
results_vs <- data_wr %>% 
  group_nest(pid) %>% 
    mutate(censored = map(data, censor, 1300, 100)) %>%
    mutate(ptb_data = map(censored, ptb)) %>%
    mutate(lifetable = map(ptb_data, setup_lt_2IV)) %>%      
    mutate(condacc = map(censored, calc_ca_2IV)) %>%     
    mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca_2IV)) %>%  
    mutate(plot = map2(.x = lifetable_ca, .y = pid, plot_eha_2IV, 1, 1))  
```

```{r}
pluck(results_vs, 8,8)
```

Make plot for subject 5, setsize = 12

Make plot for subject 8, setsize = 3


```{r}
data_search <- tibble(
  bin = results_vs[[7]][[8]] %>%
    filter(condition1 == 3 & condition2 == "Target present") %>% pull(period),
  haz_present = results_vs[[7]][[8]] %>%
    filter(condition1 == 3 & condition2 == "Target present") %>% pull(hazard),
  haz_absent = results_vs[[7]][[8]] %>%
    filter(condition1 == 3 & condition2 == "Target absent") %>% pull(hazard),
  haz_present_se = results_vs[[7]][[8]] %>%
    filter(condition1 == 3 & condition2 == "Target present") %>% pull(se_haz),
  haz_absent_se = results_vs[[7]][[8]] %>%
    filter(condition1 == 3 & condition2 == "Target absent") %>% pull(se_haz),
  ca_present = results_vs[[7]][[8]] %>%
    filter(condition1 == 3 & condition2 == "Target present") %>% pull(ca),
  ca_absent = results_vs[[7]][[8]] %>%
    filter(condition1 == 3 & condition2 == "Target absent") %>% pull(ca),
  ca_present_se = results_vs[[7]][[8]] %>%
    filter(condition1 == 3 & condition2 == "Target present") %>% pull(se_ca),
  ca_absent_se = results_vs[[7]][[8]] %>%
    filter(condition1 == 3 & condition2 == "Target absent") %>% pull(se_ca),
)
```

Plot visual search data for subj 5 and setsize 12.

```{r}
p5 <- data_search %>%
  rename(present = haz_present, absent = haz_absent) %>%
  pivot_longer(cols = present:absent,
               values_to  = "hazard"
               ) %>%
   mutate(
    se = case_when(
      name == "present" ~ haz_present_se,
      name == "absent" ~ haz_absent_se)
  ) %>%
  rename(`Target presence` = name) %>%

ggplot(aes(x=bin,y=hazard,group=`Target presence`,colour=`Target presence`)) +
    geom_linerange(aes(ymin=hazard-se, ymax=hazard+se), 
                   show.legend = F, linewidth = 0.2) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(0,seq(1,13,1)), 
                     limits = c(-0.4,11.1), 
                     labels = c(0,seq(1,13,1)*100),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)",
       y = "hazard probability",
       title = "conjunction search - setsize 12") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
         panel.grid.major.x = element_blank(),
        legend.position = "right")
p5

# plot conditional accuracy
p6 <- data_search %>%
  rename(present = ca_present, absent = ca_absent) %>%
  pivot_longer(cols = present:absent,
               values_to  = "ca"
               ) %>%
   mutate(
    se = case_when(
      name == "present" ~ ca_present_se,
      name == "absent" ~ ca_absent_se)
  ) %>%
  rename(`Target presence` = name) %>%

ggplot(aes(x=bin,y=ca,group=`Target presence`,colour=`Target presence`)) +
    geom_linerange(aes(ymin=ca-se, ymax=ca+se), 
                   show.legend = F, linewidth = 0.2) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(0,seq(1,13,1)), 
                     limits = c(-0.4,11.1), 
                     labels = c(0,seq(1,13,1)*100),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)",
       y = "conditional accuracy") +
 #      title = "conjunction search - setsize 12") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
         panel.grid.major.x = element_blank(),
        legend.position = "right")
p6
```

## Combine vs plots.


```{r}
#| label: combine-vs-plots

p5_theme <- p5 + theme(legend.position = "none")

visualsearch_plot <- (p5_theme / p6) +
    plot_layout(guides = "collect",
                axes = "collect_x") & 
    theme(legend.position = "bottom")
```

Save the plot.

```{r}
ggsave("REVIEW2/visualsearch/searchdata_plot.png", width = 10, height = 10, dpi = 800)
```


## Spatial cueing: valid and invalid

Read and wrangle data.

```{r read-raw-data-sc}
raw <- read.table("REVIEW2/spatialcueing/dataIORdiscr_central_frame.dat", header=TRUE) 
head(raw)

unique(raw$vp)
       
summary(raw)
```


```{r wrangle-data-sc}
# recode accuracy
data_sc <- raw %>%
  # remove practice trials
  filter(practice == 0) %>%
  select(-c("practice","tr","gen","age","hand",
            "bril","ss","rd","fixduration","cueduration","centralduration","void","ISI1duration","ISI2duration","targetduration","corresp","resperr","outlier")) %>%
  filter(vp < 100) %>%
  rename(pid = vp, acc = respac, soa = SOA) %>%
  mutate(rt = rt*1000)
  
  
summary(data_sc) # 6 subjects, 3 cuetypes, 2 central, 4 soa
  
head(data_sc)
summary(data_sc)

# check number of trials
data_sc %>% group_by(pid, central, cue_type, soa) %>%
  summarise(N = n())

# remove no central, add trial nr and factors for cue_type and soa
data_sc <- data_sc %>%
  filter(central == 2) %>%
  group_by(pid) %>%
  mutate(trial = 1:n(),
         cue = factor(cue_type,
                          levels = c(1,2,3),
                          labels = c("valid","invalid","no cue")),
         soa = factor(soa, levels = c(1,2,3,4),
                         labels = c("x-y","a-b","c-d","e-f"))) %>%
  mutate(condition1 = soa,
         condition2 = cue) %>%
  select(pid, trial, condition1, condition2, rt, acc)

summary(data_sc)
head(data_sc)


# add 400 ms and remove negative RTs, and select central = 2
summary(data_sc$rt)

data_sc <- data_sc %>% mutate(rt = rt + 400) %>%
  filter(rt > 0)
head(data_sc)
summary(data_sc$rt)
```

set up life tables 

```{r load-functions}
source("REVIEW2/spatialcueing/custum_functions_spatialcueing.R")
```

Apply the custom functions

```{r apply-functions}
results_sc <- data_sc %>% 
  group_nest(pid) %>% 
    mutate(censored = map(data, censor, 800, 40)) %>%
    mutate(ptb_data = map(censored, ptb)) %>%
    mutate(lifetable = map(ptb_data, setup_lt_2IV)) %>%      
    mutate(condacc = map(censored, calc_ca_2IV)) %>%     
    mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca_2IV)) %>%  
    mutate(plot = map2(.x = lifetable_ca, .y = pid, plot_eha_2IV, 1, 1))  
```

```{r}
pluck(results_sc, 8, 2)
```

Make plot for subject 2, soa=e-f or. 187 - 187 :    0 = target, -187 = central cue, -374 = peripheral cue

```{r}
data_search <- tibble(
  bin = results_sc[[7]][[2]] %>%
    filter(condition1 == "e-f" & condition2 == "valid") %>% pull(period), # 20 bins of 40 ms
  haz_valid = results_sc[[7]][[2]] %>%
    filter(condition1 == "e-f" & condition2 == "valid") %>% pull(hazard),
  haz_invalid = results_sc[[7]][[2]] %>%
    filter(condition1 == "e-f" & condition2 == "invalid") %>% pull(hazard),
  haz_valid_se = results_sc[[7]][[2]] %>%
    filter(condition1 == "e-f" & condition2 == "valid") %>% pull(se_haz),
  haz_invalid_se = results_sc[[7]][[2]] %>%
    filter(condition1 == "e-f" & condition2 == "invalid") %>% pull(se_haz),
  ca_valid = results_sc[[7]][[2]] %>%
    filter(condition1 == "e-f" & condition2 == "valid") %>% pull(ca),
  ca_invalid = results_sc[[7]][[2]] %>%
    filter(condition1 == "e-f" & condition2 == "invalid") %>% pull(ca),
  ca_valid_se = results_sc[[7]][[2]] %>%
    filter(condition1 == "e-f" & condition2 == "valid") %>% pull(se_ca),
  ca_invalid_se = results_sc[[7]][[2]] %>%
    filter(condition1 == "e-f" & condition2 == "invalid") %>% pull(se_ca),
)
```

Plot spatial cueing data for subj 2

```{r}
p7 <- data_search %>%
    rename(valid = haz_valid, invalid = haz_invalid) %>%
  pivot_longer(cols = valid:invalid,
               values_to  = "hazard"
               ) %>%
   mutate(
    se = case_when(
      name == "valid" ~ haz_valid_se,
      name == "invalid" ~ haz_invalid_se)
  ) %>%
  rename(`Cue validity` = name) %>%

ggplot(aes(x=bin,y=hazard,group=`Cue validity`,colour=`Cue validity`)) +
  
  geom_vline(xintercept = 10, linetype = 3) +
  geom_vline(xintercept = 5.5, linetype = 1) +
  geom_vline(xintercept = 1, linetype = 2) +
  
  
  geom_linerange(aes(ymin=hazard-se, ymax=hazard+se), 
                   show.legend = F, linewidth = 0.2) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(seq(1,20,1)), 
                     limits = c(-0.4,20.1), 
                     labels = c(seq(-360,400,40)),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)",
       y = "hazard probability",
       title = "spatial cueing: 2-choice (left/right) localization task") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
         panel.grid.major.x = element_blank(),
        legend.position = "right")
p8 <- p7  + annotate("text", x = 0.1, y = .5, label = "peripheral cue", angle = 90) +
  annotate("text", x = 5, y = .5, label = "central cue", angle = 90) +
  annotate("text", x = 9, y = .5, label = "peripheral target", angle = 90) 
  


# plot conditional accuracy
p9 <- data_search %>%
  rename(valid = ca_valid, invalid = ca_invalid) %>%
  pivot_longer(cols = valid:invalid,
               values_to  = "ca"
               ) %>%
   mutate(
    se = case_when(
      name == "valid" ~ ca_valid_se,
      name == "invalid" ~ ca_invalid_se)
  ) %>%
  rename(`Cue validity` = name) %>%

ggplot(aes(x=bin,y=ca,group=`Cue validity`,colour=`Cue validity`)) +
  
    geom_vline(xintercept = 10, linetype = 3) +
  geom_vline(xintercept = 5.5, linetype = 1) +
  geom_vline(xintercept = 1, linetype = 2) +
    geom_linerange(aes(ymin=ca-se, ymax=ca+se), 
                   show.legend = F, linewidth = 0.2) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(seq(1,20,1)), 
                     limits = c(-0.4,20.1), 
                     labels = c(seq(-360,400,40)),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)",
       y = "conditional accuracy") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
         panel.grid.major.x = element_blank(),
        legend.position = "right")
p10 <- p9  + annotate("text", x = 0.1, y = .5, label = "peripheral cue", angle = 90) +
  annotate("text", x = 5, y = .5, label = "central cue", angle = 90) +
  annotate("text", x = 9, y = .5, label = "peripheral target", angle = 90) 
```

## Combine ss plots.


```{r}
#| label: combine-vs-plots

p8_theme <- p8 + theme(legend.position = "none")

spatialcueing_plot <- (p8_theme / p9) +
    plot_layout(guides = "collect",
                axes = "collect_x") & 
    theme(legend.position = "bottom")
```

Save the plot.

```{r}
ggsave("REVIEW2/spatialcueing/spatialcueing_plot.png", width = 10, height = 10, dpi = 800)
```


## stop-signal: no, 80, 120, 160, 200 ms

Read and wrangle data.

```{r read-raw-data-ss}
raw <- read.table("REVIEW2/stopsignal/dataStopExp1.dat", header=T)
head(raw)
```


```{r wrangle-data-ss}
# recode accuracy
data_ss <- raw %>%
  
    # remove practice trials
  filter(practice == 0) %>%
  select(-c("practice","tr","bl","gen","age","hand",
            "bril","ss","rd","fixduration","goduration","SSDduration","void","stopduration","corresp","resperr","outlier")) %>%
  filter(vp < 100) %>%
  rename(pid = vp, acc = respac, ssd = SSD, ss = Stop_signal) %>%
  mutate(rt = rt*1000,
         condition = ifelse(ss == 2, 5,
                            ifelse(ss==1 & ssd==1,1,
                            ifelse(ss==1 & ssd==2,2,
                            ifelse(ss==1 & ssd==3,3,4))))) %>%
  filter(rt > 0)
  

head(data_ss)
summary(data_ss) # 5 participants, 5 conditions

# check number of trials
data_ss %>% group_by(pid, condition) %>%
  summarise(N = n()) %>% print(n=12)

# add trial nr and factors for target-presence and set-size
data_ss <- data_ss %>%
  group_by(pid) %>%
  mutate(trial = 1:n(),
         condition = factor(condition,
                          levels = c(1,2,3,4,5),
                          labels = c("ssd = 80 ms","ssd = 120 ms","ssd = 160 ms","ssd = 200 ms","no stop signal"))) %>%
  select(pid, trial, condition, rt, acc)

summary(data_ss)
head(data_ss)
```

set up life tables 

```{r load-functions-vs}
source("REVIEW2/stopsignal/custom_functions_stopsignal.R")
```

Apply the custom functions

```{r apply-functions-vs}
results_ss <- data_ss %>% 
  group_nest(pid) %>% 
    mutate(censored = map(data, censor, 500, 20)) %>%
    mutate(ptb_data = map(censored, ptb)) %>%
    mutate(lifetable = map(ptb_data, setup_lt)) %>%      
    mutate(condacc = map(censored, calc_ca)) %>%     
    mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%  
    mutate(plot = map2(.x = lifetable_ca, .y = pid, plot_eha, 1, 1))  
```

```{r}
pluck(results_ss, 8,1)
```

Make plot for subject 1


```{r}
data_stop <- tibble(
  bin = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 80 ms") %>% pull(period),
  haz_no = results_vs[[7]][[1]] %>%
    filter(condition == "no stop signal") %>% pull(hazard),
  haz_80 = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 80 ms") %>% pull(hazard),
  haz_120 = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 120 ms") %>% pull(hazard),
  haz_160 = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 160 ms") %>% pull(hazard),
  haz_200 = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 200 ms") %>% pull(hazard),
  haz_80se = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 80 ms") %>% pull(se_haz),
  haz_120se = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 120 ms") %>% pull(se_haz),
  haz_160se = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 160 ms") %>% pull(se_haz),
  haz_200se = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 200 ms") %>% pull(se_haz),
  haz_nose = results_vs[[7]][[1]] %>%
    filter(condition == "no stop signal") %>% pull(se_haz),
  
  ca_no = results_vs[[7]][[1]] %>%
    filter(condition == "no stop signal") %>% pull(ca),
  ca_80 = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 80 ms") %>% pull(ca),
  ca_120 = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 120 ms") %>% pull(ca),
  ca_160 = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 160 ms") %>% pull(ca),
  ca_200 = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 200 ms") %>% pull(ca),
  ca_80se = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 80 ms") %>% pull(se_ca),
  ca_120se = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 120 ms") %>% pull(se_ca),
  ca_160se = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 160 ms") %>% pull(se_ca),
  ca_200se = results_vs[[7]][[1]] %>%
    filter(condition == "ssd = 200 ms") %>% pull(se_ca),
  ca_nose = results_vs[[7]][[1]] %>%
    filter(condition == "no stop signal") %>% pull(se_ca)

)

tail(data_stop)
```

Plot stop signal data.

```{r}


# Color deficient safe palette with grey and black:
cbPalette <- c(
  "#000000", "#56B4E9","#E69F00","#F0E442","#CC79A7",
  "#999999",   "#009E73", 
  "#0072B2", "#D55E00"
)


p11 <- data_stop %>%
  #rename(present = haz_present, absent = haz_absent) %>%
  pivot_longer(cols = haz_no:haz_200,
               values_to  = "hazard"
               ) %>%
   mutate(
    se = case_when(
      name == "haz_no" ~ haz_nose,
      name == "haz_80" ~ haz_80se,
      name == "haz_120" ~ haz_120se,
      name == "haz_160" ~ haz_160se,
      name == "haz_200" ~ haz_200se)
  ) %>%
  rename(`Trial type` = name) %>%
  mutate(`Trial type` = factor(`Trial type`, 
                               levels = c("haz_no","haz_80","haz_120","haz_160","haz_200"),
                               labels = c("No stop signal", "ssd = 80 ms", "ssd = 120 ms", "ssd = 160 ms", "ssd = 200 ms"))) %>%
  
  filter(bin > 3) %>%

ggplot(aes(x=bin,y=hazard,group=`Trial type`,colour=`Trial type`)) +
  
  geom_vline(xintercept = 4, linetype = 1, color = "#56B4E9") +
  geom_vline(xintercept = 6, linetype = 1, color = "#E69F00") +
  geom_vline(xintercept = 8, linetype = 1, color = "#F0E442") +
  geom_vline(xintercept = 10, linetype = 1, color = "#CC79A7") +
  
  
    geom_linerange(aes(ymin=hazard-se, ymax=hazard+se), 
                   show.legend = F, linewidth = 0.2) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
  scale_color_manual(values = cbPalette) +
  scale_y_continuous(limits = c(0,.501), 
                     breaks = c(0,.25,.5),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(seq(4,25,1)), 
                     limits = c(3.6,25.1), 
                     labels = c(seq(4,25,1)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)",
       y = "hazard probability",
       title = "stop-signal task") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
         panel.grid.major.x = element_blank(),
        legend.position = "bottom") + guides(colour=guide_legend(nrow=2, byrow=TRUE))
p11

# plot conditional accuracy
p12 <- data_stop %>%
  #rename(present = ca_present, absent = ca_absent) %>%
  pivot_longer(cols = ca_no:ca_200,
               values_to  = "ca"
               ) %>%
   mutate(
    se = case_when(
      name == "ca_no" ~ ca_nose,
      name == "ca_80" ~ ca_80se,
      name == "ca_120" ~ ca_120se,
      name == "ca_160" ~ ca_160se,
      name == "ca_200" ~ ca_200se)
  ) %>%
  rename(`Trial type` = name) %>%
   mutate(`Trial type` = factor(`Trial type`, 
                               levels = c("ca_no","ca_80","ca_120","ca_160","ca_200"),
                               labels = c("No stop signal", "ssd = 80 ms", "ssd = 120 ms", "ssd = 160 ms", "ssd = 200 ms"))) %>%
  
    filter(bin > 3) %>%

ggplot(aes(x=bin,y=ca,group=`Trial type`,colour=`Trial type`)) +
    geom_vline(xintercept = 4, linetype = 1, color = "#56B4E9") +
  geom_vline(xintercept = 6, linetype = 1, color = "#E69F00") +
  geom_vline(xintercept = 8, linetype = 1, color = "#F0E442") +
  geom_vline(xintercept = 10, linetype = 1, color = "#CC79A7") +
  
    geom_linerange(aes(ymin=ca-se, ymax=ca+se), 
                   show.legend = F, linewidth = 0.2) +
  geom_line() + geom_point(size = .5) +
  theme_bw() +
    scale_color_manual(values = cbPalette) +
  scale_y_continuous(limits = c(0,1.001), 
                     breaks = c(0,.25,.5,.75,1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_x_continuous(breaks = c(seq(4,25,1)), 
                     limits = c(3.6,25.1), 
                     labels = c(seq(4,25,1)*20),
                     expand = expansion(mult = c(0, 0))) +
  labs(x = "time bin (upper limit in ms)",
       y = "conditional accuracy") +
 #      title = "conjunction search - setsize 12") +
  theme(axis.text.x = element_text(angle=90),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
         panel.grid.major.x = element_blank(),
        legend.position = "bottom") + guides(colour=guide_legend(nrow=2, byrow=TRUE))
p12
```

## Combine stop signal plots.


```{r}
#| label: combine-vs-plots

p11_theme <- p11 + theme(legend.position = "none")

stopsignal_plot <- (p11_theme / p12) +
    plot_layout(guides = "collect",
                axes = "collect_x") & 
    theme(legend.position = "bottom")
```

Save the plot.

```{r}
ggsave("REVIEW2/visualsearch/searchdata_plot.png", width = 10, height = 10, dpi = 800)
```

## Combine into fig 5

```{r}

(visualsearch_plot|spatialcueing_plot|stopsignal_plot) +
    plot_layout(axes = "collect_y") & 
    theme(legend.position = "bottom")

```

Save figure 5.

```{r}
ggsave("REVIEW2/Plot_distributions_empirical.png", width = 14, height = 10, dpi = 800)
```


# Figure ...




